### 2023-08-07 docs/paper.tex
Add unfinished paper report

\documentclass[conference]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Implementing XBF: An Efficient Self-Describing Binary Format}

\author{
	\IEEEauthorblockN{David Krauthamer}
	\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
		\textit{name of organization (of Aff.)}\\
		City, Country \\
		dkrautha@pm.me}
	\and
	\IEEEauthorblockN{Dov Kruger}
	\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
		\textit{name of organization (of Aff.)}\\
		City, Country \\
		email address or ORCID}
	% \and
	% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
	% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
	% 	\textit{name of organization (of Aff.)}\\
	% 	City, Country \\
	% 	email address or ORCID}
}

\maketitle

\begin{abstract}
	Do later
\end{abstract}

\begin{IEEEkeywords}
	Data format, XML, JSON, Binary data
\end{IEEEkeywords}

\section{Introduction}

On the web, data is largely interchanged in one of two formats, XML (Extensible Markup Language)\cite{xml} or JSON (JavaScript Object Notation)\cite{json}. These formats have two key features that have led to their popularity, human-readability and being self-describing. These features have inherent inefficiencies, which XBF (Xtensible Binary Format)\cite{xbf_spec} intends to solve by making compromises.

\subsection{Self-Description}

A self-describing format is a format in which a recipient is not required to have a schema for the incoming data, and the names and types of the data being sent can be determined from the data itself. The following is an example of XML:

\begin{lstlisting}[language=XML]
<person>
	<name>John Jackson</name>
	<age>25</age>
</person>
\end{lstlisting}

This example represents a person object with two fields, a string name and integer age. The downside to being self-describing is wasted characters on the metadata describing the object. In this particular example, 14 characters of the 59 total characters actually contain the desired data, the rest are just describing the metadata. The best thing a format can do is minimize this overhead as much as possible, it cannot be eliminated while still being self-describing.

An additional optimization that could be made (but isn't with either of these formats) is deduplication of the metadata. Both JSON and XML include the metadata for every element sent, even if what is being sent is a homogenous list of elements that all share the same structure.

\subsection{Human Readability}

A human-readable format is one where data is encoded entirely as ASCII or UTF-8 characters, and formatted in such a way that a person could easily read or write it without the need of a machine. Human-readability brings with it two inefficiencies, the need to parse text data into a binary format that a computer understands, and additional overhead from whitespace characters.

\subsection{Enhancements to be Made}

With the inefficiencies of these two formats in mind, XBF seeks to improve on them through a few core design decisions:

\begin{IEEEitemize}
	\item Send data and metadata in a binary format (no longer human-readable)
	\item Remain self-describing, but have the option to not include metadata if it's already known
	\item Prioritize simplicity in the type system
\end{IEEEitemize}

% \begin{table}[htbp]
% 	\caption{Table Type Styles}
% 	\begin{center}
% 		\begin{tabular}{|c|c|c|c|}
% 			\hline
% 			\textbf{Table} & \multicolumn{3}{|c|}{\textbf{Table Column Head}}                                                         \\
% 			\cline{2-4}
% 			\textbf{Head}  & \textbf{\textit{Table column subhead}}           & \textbf{\textit{Subhead}} & \textbf{\textit{Subhead}} \\
% 			\hline
% 			copy           & More table copy$^{\mathrm{a}}$                   &                           &                           \\
% 			\hline
% 			\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
% 		\end{tabular}
% 		\label{tab1}
% 	\end{center}
% \end{table}

\section*{Acknowledgment}

\bibliographystyle{IEEEtran}
\bibliography{refs}
\end{document}

### 2023-08-08 xbf_rs/examples/format_comparison_client/main.rs
Use date and volume fields from sony data
Removes random string heavy test

@@ -64,40 +64,36 @@ fn main() -> Result<()> {
         std::env::args().nth(1).unwrap_or("127.0.0.1".to_string())
     );
 
-    for request_type in 0..RequestType::Unknown.into() {
-        for data_format in 0..DataFormat::Unknown.into() {
-            let mut durations = vec![];
-            let mut bytes_read = None;
+    for data_format in 0..DataFormat::Unknown.into() {
+        let mut durations = vec![];
+        let mut bytes_read = None;
 
-            for _ in 0..100 {
-                let time_start = Instant::now();
+        for _ in 0..100 {
+            let time_start = Instant::now();
 
-                let mut connection = TcpStream::connect(&addr)?;
-                connection.write_u8(request_type)?;
-                connection.write_u8(data_format)?;
+            let mut connection = TcpStream::connect(&addr)?;
+            connection.write_u8(data_format)?;
 
-                let mut buf = vec![];
-                let bytes = connection.read_to_end(&mut buf)?;
+            let mut buf = vec![];
+            let bytes = connection.read_to_end(&mut buf)?;
 
-                let time_elapsed = Instant::now() - time_start;
+            let time_elapsed = Instant::now() - time_start;
 
-                if let None = bytes_read {
-                    bytes_read = Some(bytes);
-                }
-
-                durations.push(time_elapsed);
+            if let None = bytes_read {
+                bytes_read = Some(bytes);
             }
 
-            let avg_time = durations.iter().sum::<std::time::Duration>() / durations.len() as u32;
-
-            println!(
-                "Request Type: {:?}\nData Format: {:?}\nAvg Time: {:?}\nBytes Read: {:?}\n",
-                RequestType::from(request_type),
-                DataFormat::from(data_format),
-                avg_time,
-                bytes_read
-            );
+            durations.push(time_elapsed);
         }
+
+        let avg_time = durations.iter().sum::<std::time::Duration>() / durations.len() as u32;
+
+        println!(
+            "Data Format: {:?}\nAvg Time: {:?}\nBytes Read: {:?}\n",
+            DataFormat::from(data_format),
+            avg_time,
+            bytes_read
+        );
     }
 
     Ok(())

### 2023-08-08 xbf_rs/examples/format_comparison_server/main.rs

@@ -1,4 +1,3 @@
-use fake::{faker::address::en::StreetName, faker::name::en::Name, Dummy};
 use reqwest::Client;
 use serde::{Deserialize, Serialize};
 use std::sync::Arc;
@@ -13,6 +12,8 @@ use xbf_rs::{
 
 #[derive(Serialize, Deserialize)]
 struct StockRecord {
+    #[serde(rename = "Date")]
+    date: String,
     #[serde(rename = "Open")]
     open: f64,
     #[serde(rename = "High")]
@@ -23,16 +24,8 @@ struct StockRecord {
     close: f64,
     #[serde(rename(serialize = "AdjClose", deserialize = "Adj Close"))]
     adj_close: f64,
-}
-
-#[derive(Serialize, Dummy)]
-struct Person {
-    #[dummy(faker = "Name()")]
-    name: String,
-    #[dummy(faker = "1..120")]
-    age: u8,
-    #[dummy(faker = "StreetName()")]
-    address: String,
+    #[serde(rename = "Volume")]
+    volume: u64,
 }
 
 async fn get_yahoo_data() -> Result<String, reqwest::Error> {
@@ -51,7 +44,9 @@ fn get_native_vec_from_csv(csv_content: &str) -> Result<Vec<StockRecord>, csv::E
 
 fn to_csv<T: Serialize>(records: &[T]) -> anyhow::Result<Vec<u8>> {
     let mut csv_writer = csv::Writer::from_writer(vec![]);
-    csv_writer.serialize(records)?;
+    records
+        .iter()
+        .try_for_each(|record| csv_writer.serialize(record))?;
     Ok(csv_writer.into_inner()?)
 }
 
@@ -77,11 +72,13 @@ fn stocks_to_xbf(records: &[StockRecord]) -> Result<Vec<u8>, std::io::Error> {
     let sr_xbf_metadata = XbfStructMetadata::new(
         "StockRecord",
         indexmap::indexmap! {
+            "Date" => XbfPrimitiveMetadata::String.into(),
             "Open" => XbfPrimitiveMetadata::F64.into(),
             "High" => XbfPrimitiveMetadata::F64.into(),
             "Low" => XbfPrimitiveMetadata::F64.into(),
             "Close" => XbfPrimitiveMetadata::F64.into(),
             "AdjClose" => XbfPrimitiveMetadata::F64.into(),
+            "Volume" => XbfPrimitiveMetadata::U64.into(),
         },
     );
 
[…]

### 2023-08-08 docs/paper.tex
Add latest paper changes to ensure they aren't lost

@@
 
-On the web, data is largely interchanged in one of two formats, XML (Extensible Markup Language)\cite{xml} or JSON (JavaScript Object Notation)\cite{json}.
+On the web, data is largely interchanged in one of two formats, XML (Extensible Markup Language)\cite{xml_spec} or JSON (JavaScript Object Notation)\cite{json_spec}.
 These formats have two key features that have led to their popularity, human-readability and being self-describing.
-These features have inherent inefficiencies, which XBF (Xtensible Binary Format)\cite{xbf_spec} intends to solve by making compromises.
+These features have inherent inefficiencies, which XBF (Xtensible Binary Format)\cite{xbf_spec} intends to solve through carefully chosen compromises.
 
@@
 
-\subsection{Enhancements to be Made}
+\subsection{Enhancements Made}
 
@@
 \begin{IEEEitemize}
-	\item Send data and metadata in a binary format (no longer human-readable)
-	\item Remain self-describing, but have the option to not include metadata if it's already known
-	\item Prioritize simplicity in the type system
+	\item Send data and metadata in a binary format (no longer human-readable).
+	\item Remain self-describing, but have the option to not include metadata if it's already known, and deduplicate metadata whenever possible.
+	\item Prioritize simplicity in the type system.
 \end{IEEEitemize}
+
+\section{Methodology and Implementation}
+
+\section{Results}
+
+\section{Discussion}
+
+\section{Conclusion}
+
+\section{}
 
@@
 
-\section*{Acknowledgment}
-
 \bibliographystyle{IEEEtran}

### 2023-08-08 docs/paper.tex
Update paper

@@
 
-\section{Methodology and Implementation}
+\section{Format Design}
+
+\section{Implementation}
+
+\section{Evaluation}
+
+XBF was evaluated against a number of other self-describing data interchange formats.
+
+\begin{IEEEitemize}
+	\item CSV (Comma Separated Values) \cite{csv_spec}\cite{csv_parser}
+	\item MessagePack \cite{msgpack_spec}\cite{msgpack_parser}
+	\item CBOR (Concise Binary Object Representation) \cite{cbor_spec}\cite{cbor_parser}
+	\item JSON \cite{json_parser}
+	\item XML \cite{xml_parser}
+\end{IEEEitemize}
+
+The primary goal of the test was to measure how many bytes a particular format required to transmit a given set of data, with the secondary objective being to measure the timed performance of the format.
+The data\cite{sony_stock_data} was a year of.
 

### 2023-08-08 docs/paper.tex
Update paper

@@
 \usepackage{textcomp}
+\usepackage{xurl}
 \usepackage{xcolor}
@@
 \begin{IEEEitemize}
-	\item CSV (Comma Separated Values) \cite{csv_spec}\cite{csv_parser}
-	\item MessagePack \cite{msgpack_spec}\cite{msgpack_parser}
-	\item CBOR (Concise Binary Object Representation) \cite{cbor_spec}\cite{cbor_parser}
-	\item JSON \cite{json_parser}
-	\item XML \cite{xml_parser}
+	\item CSV (Comma Separated Values)\cite{csv_spec}\cite{csv_parser}
+	\item MessagePack\cite{msgpack_spec}\cite{msgpack_parser}
+	\item CBOR (Concise Binary Object Representation)\cite{cbor_spec}\cite{cbor_parser}
+	\item JSON\cite{json_parser}
+	\item XML\cite{xml_parser}
 \end{IEEEitemize}

### 2023-08-09 xbf_rs/examples/format_comparison_client/main.rs
Add data only for xbf

@@ -4,30 +4,6 @@ use std::{io::Read, net::TcpStream, time::Instant};
 
 #[repr(u8)]
 #[derive(Debug)]
-enum RequestType {
-    Stock,
-    Person,
-    Unknown,
-}
-
-impl From<u8> for RequestType {
-    fn from(value: u8) -> Self {
-        match value {
-            0 => Self::Stock,
-            1 => Self::Person,
-            _ => Self::Unknown,
-        }
-    }
-}
-
-impl From<RequestType> for u8 {
-    fn from(value: RequestType) -> Self {
-        value as u8
-    }
-}
-
-#[repr(u8)]
-#[derive(Debug)]
 enum DataFormat {
     Csv,
     MessagePack,
@@ -35,6 +11,7 @@ enum DataFormat {
     Json,
     Xml,
     Xbf,
+    XbfDataOnly,
     Unknown,
 }
 
@@ -47,6 +24,7 @@ impl From<u8> for DataFormat {
             3 => Self::Json,
             4 => Self::Xml,
             5 => Self::Xbf,
+            6 => Self::XbfDataOnly,
             _ => Self::Unknown,
         }
     }
@@ -79,7 +57,7 @@ fn main() -> Result<()> {
 
             let time_elapsed = Instant::now() - time_start;
 
-            if let None = bytes_read {
+            if bytes_read.is_none() {
                 bytes_read = Some(bytes);
             }
 

### 2023-08-09 xbf_rs/examples/format_comparison_server/main.rs

@@ -68,7 +68,7 @@ fn to_xml<T: Serialize>(records: &[T]) -> Result<Vec<u8>, quick_xml::de::DeError
     Ok(quick_xml::se::to_string_with_root("root", records)?.into_bytes())
 }
 
-fn stocks_to_xbf(records: &[StockRecord]) -> Result<Vec<u8>, std::io::Error> {
+fn create_xbf_vec(records: &[StockRecord]) -> XbfVec {
     let sr_xbf_metadata = XbfStructMetadata::new(
         "StockRecord",
         indexmap::indexmap! {
@@ -82,9 +82,9 @@ fn stocks_to_xbf(records: &[StockRecord]) -> Result<Vec<u8>, std::io::Error> {
         },
     );
 
-    let vec = XbfVec::new_unchecked(
+    XbfVec::new_unchecked(
         XbfVecMetadata::new(sr_xbf_metadata.clone()),
-        records.into_iter().map(|record| {
+        records.iter().map(|record| {
             XbfStruct::new_unchecked(
                 sr_xbf_metadata.clone(),
                 [
@@ -97,8 +97,19 @@ fn stocks_to_xbf(records: &[StockRecord]) -> Result<Vec<u8>, std::io::Error> {
                 ],
             )
         }),
-    );
+    )
+}
 
+fn to_xbf(records: &[StockRecord]) -> Result<Vec<u8>, std::io::Error> {
+    let vec = create_xbf_vec(records);
+    let mut bytes = vec![];
+    vec.serialize_vec_type(&mut bytes)?;
+
+    Ok(bytes)
+}
+
+fn to_xbf_data_only(records: &[StockRecord]) -> Result<Vec<u8>, std::io::Error> {
+    let vec = create_xbf_vec(records);
     let mut bytes = vec![];
     vec.get_metadata().serialize_vec_metadata(&mut bytes)?;
     vec.serialize_vec_type(&mut bytes)?;
@@ -115,6 +126,7 @@ enum DataFormat {
     Json,
     Xml,
     Xbf,
+    XbfDataOnly,
     Unknown,
 }
 
@@ -127,6 +139,7 @@ impl From<u8> for DataFormat {
             3 => Self::Json,
             4 => Self::Xml,
             5 => Self::Xbf,
+            6 => Self::XbfDataOnly,
             _ => Self::Unknown,
         }
     }
@@ -145,7 +158,8 @@ async fn handle_stock_request(
         DataFormat::Cbor => to_cbor(records)?,
[…]

### 2023-08-09 docs/paper.tex
Update paper and references

@@
 
-The primary goal of the test was to measure how many bytes a particular format required to transmit a given set of data, with the secondary objective being to measure the timed performance of the format.
-The data\cite{sony_stock_data} was a year of.
+The test consisted of two components, a client and a server, both written in Rust.
+The server downloads a year of Sony stock data from Yahoo Finance\cite{sony_stock_data} in CSV format, and parses it into a list of objects.
+The original size of the CSV and a calculated value for how much memory is taken up by the parsed list is logged.
+It then waits for connections, and depending on the request received, serializes the list of objects into the requested format and sends it to the client.
+The serialized data is not cached, and the serialization is performed for each request.
+
+The client performs the time measurement, as well as initiates connections to the server.
+For each data format, the client sends 100 requests to the server, recording the time it takes from initiating the connection to receiving all the data.
+The average of the time for these 100 requests is then recorded, along with how large each response received was in bytes.
 
 \section{Results}
+
+\begin{table}[htbp]
+	\begin{center}
+		\begin{tabular}{|c|c|c|}
+			\hline
+			\textbf{Format} & \textbf{Avg Time (ms)} & \textbf{Bytes Read} \\
+			\hline
+			CSV             & 18.931802              & 16411               \\
+			MessagePack     & 11.220957              & 15565               \\
+			CBOR            & 16.957873              & 25507               \\
+			JSON            & 21.912745              & 31180               \\
+			XML             & 21.873043              & 43699               \\
+			XBF             & 11.322245              & 14686               \\
+			\hline
+		\end{tabular}
+		\label{results}
+	\end{center}
+\end{table}
+
+See the \ref{results}
 
@@
 
-\section{}
-
-% \begin{table}[htbp]
-% 	\caption{Table Type Styles}
-% 	\begin{center}
-% 		\begin{tabular}{|c|c|c|c|}
-% 			\hline
-% 			\textbf{Table} & \multicolumn{3}{|c|}{\textbf{Table Column Head}}                                                         \\
-% 			\cline{2-4}
-% 			\textbf{Head}  & \textbf{\textit{Table column subhead}}           & \textbf{\textit{Subhead}} & \textbf{\textit{Subhead}} \\
-% 			\hline
-% 			copy           & More table copy$^{\mathrm{a}}$                   &                           &                           \\
-% 			\hline
-% 			\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-% 		\end{tabular}
-% 		\label{tab1}
-% 	\end{center}
-% \end{table}
-
 \bibliographystyle{IEEEtran}

### 2023-08-09 docs/paper.tex
Finish results section

@@
 
+Should this section be here / how in depth should this go? Should this mainly just be a reference to the spec / the paper written by Dov, or should more be spoken about here?
+
 \section{Implementation}
+
+The reference implementation for XBF\cite{xbf_impl} was written in Rust. Rust was chosen for the following reasons:
+\begin{IEEEitemize}
+	\item Rust is statically typed and compiled, and allows for similarly efficient code generation to C/C++.
+	\item A powerful type system with tagged unions built in such that dynamic dispatch and virtual method tables weren't required.
+	\item Control over the allocation and de-allocation of memory.
+	\item Rust provides statically checked memory safety.
+\end{IEEEitemize}
 
@@
 The test consisted of two components, a client and a server, both written in Rust.
-The server downloads a year of Sony stock data from Yahoo Finance\cite{sony_stock_data} in CSV format, and parses it into a list of objects.
-The original size of the CSV and a calculated value for how much memory is taken up by the parsed list is logged.
-It then waits for connections, and depending on the request received, serializes the list of objects into the requested format and sends it to the client.
-The serialized data is not cached, and the serialization is performed for each request.
+The server downloaded a year of Sony stock data from Yahoo Finance\cite{sony_stock_data} in CSV format, and parsed it into a list of objects.
+Next, the original size of the CSV and a calculated value for how much memory is taken up by the parsed list was logged.
+Finally, the server waited for connections, and depending on the request received, serialized the list of objects into the requested format and sent it to the client.
+The serialized data was not cached, and the serialization was performed for each request.
 
-The client performs the time measurement, as well as initiates connections to the server.
-For each data format, the client sends 100 requests to the server, recording the time it takes from initiating the connection to receiving all the data.
-The average of the time for these 100 requests is then recorded, along with how large each response received was in bytes.
+The client performed the time measurement, as well as initiated connections to the server.
+For each data format, the client sent 100 requests to the server, and recorded the time it took from initiating the connection to receiving all the data.
+The average of the time for these 100 requests was then recorded, along with how large each response received was in bytes.
+
+The server was a DigitalOcean Droplet\cite{digital_ocean} running Ubuntu 20.04.5 located in New York City. The client was a laptop running openSUSE Tumbleweed located in Hoboken, NJ.
 
@@
 		\end{tabular}
-		\label{results}
-	\end{center}
+	\end{center}\label{time_and_bytes}
 \end{table}
 
-See the \ref{results}
+\begin{table}[htbp]
+	\begin{center}
+		\begin{tabular}{|c|c|c|}
+			\hline
+			\textbf{Format} & \textbf{Overhead (bytes)} & \textbf{Percent Overhead} \\
+			\hline
+			CSV             & 1823                      & 11.1084                   \\
+			MessagePack     & 977                       & 6.2769                    \\
+			CBOR            & 10919                     & 42.8079                   \\
+			JSON            & 16592                     & 53.2136                   \\
+			XML             & 29111                     & 66.6171                   \\
+			XBF             & 98                        & 0.6673                    \\
+			\hline
+		\end{tabular}
+	\end{center}\label{overhead}
+\end{table}
+
+Original CSV data size recorded by the server: 17160 bytes.
+
+Native data size recorded by the server: 14558 bytes.
 

### 2023-08-09 docs/paper.tex
Update paper

@@
 
+Similar to CSV (cite here?), we send a single header up front, followed by all of the data.
+For a primitive type there isn't much savings here, but there are massive savings with vectors and structs.
+With a vector the metadata is a discriminant indicating a vector is being sent, which is then follwed by the metadata for its internal type.
+This process continues recursively if the inner type (or any of its inner types) contains a vector or struct, ensuring that all metadata necessary to represent the data coming afterwards is received up front.
+The data is then sent in a single continuous block, something after this...
+
 \section{Implementation}
@@
 \end{IEEEitemize}
+
+\subsection{Primitives}
+
+The implementation of primitives is a tagged union of all possible primitive types, mapping the XBF type to the Rust native equivalent as shown in Table \ref{type_map}.
+The exceptions are the 256 bit integer types, which do not have an analog, and are instead represented by an array of four unsigned 64 bit integers.
+
+Metadata for integers was an enumeration of the possible metadata discriminants, ranging from 0 to 16, represented as unsigned 8 bit integers.
+
+\begin{table}[htbp]
+	\caption{XBF Type Map}\label{type_map}
+	\begin{center}
+		\begin{tabular}{|c|c|}
+			\hline
+			\textbf{XBF Type} & \textbf{Rust Type}          \\
+			\hline
+			Boolean           & bool                        \\
+			U8                & u8                          \\
+			U16               & u16                         \\
+			U32               & u32                         \\
+			U64               & u64                         \\
+			U128              & u128                        \\
+			U256              & [u64; 4]                    \\
+			I8                & i8                          \\
+			I16               & i16                         \\
+			I32               & i32                         \\
+			I64               & i64                         \\
+			I128              & i128                        \\
+			I256              & [u64; 4]                    \\
+			F32               & f32                         \\
+			F64               & f64                         \\
+			Bytes             & Vec\textless u8\textgreater \\
+			String            & String                      \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+\subsection{Vectors}
+
+\subsection{Structs}
 
@@
 \begin{table}[htbp]
+	\caption{Average Time and Bytes Read}
 	\begin{center}
@@
 \begin{table}[htbp]
+	\caption{Format Overheads}
 	\begin{center}

### 2023-08-10 docs/paper.tex
Add microtype package
Update author section

@@
 \usepackage{listings}
+\usepackage{microtype}
 
@@
 	\IEEEauthorblockN{David Krauthamer}
-	\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-		\textit{name of organization (of Aff.)}\\
-		City, Country \\
+	\IEEEauthorblockA{\textit{Electrical and Computer Engineering Department} \\
+		\textit{Stevens Institute of Technology}\\
+		Hoboken, USA \\
 		dkrautha@pm.me}
@@
 	\IEEEauthorblockN{Dov Kruger}
-	\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-		\textit{name of organization (of Aff.)}\\
-		City, Country \\
-		email address or ORCID}
-	% \and
-	% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-	% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-	% 	\textit{name of organization (of Aff.)}\\
-	% 	City, Country \\
-	% 	email address or ORCID}
+	\IEEEauthorblockA{\textit{Electrical and Computer Engineering Department} \\
+		\textit{Stevens Institute of Technology}\\
+		Hoboken, USA \\
+		dkruger@stevens.edu}
 }

### 2023-08-10 docs/paper.tex
Update references and other stuff

@@
 
-Should this section be here / how in depth should this go? Should this mainly just be a reference to the spec / the paper written by Dov, or should more be spoken about here?
-
-Similar to CSV (cite here?), we send a single header up front, followed by all of the data.
+Similar to CSV (cite here?), we send a single header up front, followed by all the data.
 For a primitive type there isn't much savings here, but there are massive savings with vectors and structs.
-With a vector the metadata is a discriminant indicating a vector is being sent, which is then follwed by the metadata for its internal type.
-This process continues recursively if the inner type (or any of its inner types) contains a vector or struct, ensuring that all metadata necessary to represent the data coming afterwards is received up front.
+With a vector the metadata is a discriminant indicating a vector is being sent, which is then followed by the metadata for its internal type.
+This process continues recursively if the inner type (or any of its inner types) contains a vector or struct, ensuring that all metadata necessary to represent the data coming afterward is received up front.
 The data is then sent in a single continuous block, something after this...
@@
 
-The reference implementation for XBF\cite{xbf_impl} was written in Rust. Rust was chosen for the following reasons:
+The reference implementation for XBF\cite{xbf_impl} was written in Rust\cite{rust}. Rust was chosen for the following reasons:
 \begin{IEEEitemize}
 	\item Rust is statically typed and compiled, and allows for similarly efficient code generation to C/C++.
-	\item A powerful type system with tagged unions built in such that dynamic dispatch and virtual method tables weren't required.
-	\item Control over the allocation and de-allocation of memory.
-	\item Rust provides statically checked memory safety.
+	\item A powerful type system with tagged unions built-in\cite{enums} such that dynamic dispatch and virtual method tables weren't required, and structural pattern matching was used instead.
+	\item Rust provides statically checked memory safety through the concept of ownership\cite{ownership}.
+	\item A built-in testing system in the standard library\cite{testing}.
 \end{IEEEitemize}
 
-\subsection{Primitives}
+\subsection{Values}
+
+At the top of the hierarchy of types is a generic value, named the XbfType. This is a tagged union of all possible types, consisting of Primitives (XbfPrimitive) \ref{primitives}, Vectors (XbfVec)
+\ref{vec}, and Structs (XbfStruct) \ref{struct}.
+The corresponding metadata is the XbfMetadata, a tagged union of metadata types of each of the aforementioned subtypes.
+In order to perform a downcast on either of these top level types, Rust's structural pattern matching must be used.
+
+\subsection{XbfPrimitive}\label{primitives}
 
 The implementation of primitives is a tagged union of all possible primitive types, mapping the XBF type to the Rust native equivalent as shown in Table \ref{type_map}.
-The exceptions are the 256 bit integer types, which do not have an analog, and are instead represented by an array of four unsigned 64 bit integers.
-
-Metadata for integers was an enumeration of the possible metadata discriminants, ranging from 0 to 16, represented as unsigned 8 bit integers.
+The exceptions are the 256-bit integer types, which do not have an analog, and are instead represented by an array of four unsigned 64-bit integers.
 
@@
 	\begin{center}
-		\begin{tabular}{|c|c|}
+		\begin{tabular}{|c|c|c|}
 			\hline
-			\textbf{XBF Type} & \textbf{Rust Type}          \\
+			\textbf{XBF Type} & \textbf{Rust Type}          & \textbf{C++ Type}                   \\
 			\hline
-			Boolean           & bool                        \\
-			U8                & u8                          \\
-			U16               & u16                         \\
-			U32               & u32                         \\
-			U64               & u64                         \\
-			U128              & u128                        \\
-			U256              & [u64; 4]                    \\
-			I8                & i8                          \\
-			I16               & i16                         \\
-			I32               & i32                         \\
-			I64               & i64                         \\
-			I128              & i128                        \\
-			I256              & [u64; 4]                    \\
-			F32               & f32                         \\
-			F64               & f64                         \\
-			Bytes             & Vec\textless u8\textgreater \\
-			String            & String                      \\
+			Boolean           & bool                        & bool                                \\
+			U8                & u8                          & uint8\_t                            \\
+			U16               & u16                         & uint16\_t                           \\
+			U32               & u32                         & uint32\_t                           \\
+			U64               & u64                         & uint64\_t                           \\
+			U128              & u128                        & uint64\_t[2]                        \\
+			U256              & [u64; 4]                    & uint64\_t[4]                        \\
+			I8                & i8                          & int8\_t                             \\
+			I16               & i16                         & int16\_t                            \\
+			I32               & i32                         & int32\_t                            \\
+			I64               & i64                         & int64\_t                            \\
+			I128              & i128                        & uint64\_t[2]                        \\
+			I256              & [u64; 4]                    & uint64\_t[4]                        \\
+			F32               & f32                         & float                               \\
+			F64               & f64                         & double                              \\
+			Bytes             & Vec\textless u8\textgreater & std::vector\textless u8\textgreater \\
+			String            & String                      & std::string                         \\
 			\hline
@@
 
-\subsection{Vectors}
+Metadata for integers was an enumeration of the possible metadata discriminant values, ranging from 0 to 16, represented as unsigned 8-bit integers.
 
-\subsection{Structs}
+\subsection{XbfVec}\label{vec}
+
+XbfVec was implemented as native dynamic array of XbfType values, and a reference counted pointer to the metadata of its inner type.
+In order to ensure XbfVec remains homogenous, it requires the metadata to be passed in at construction time.
+
+\subsection{XbfStruct}\label{struct}
 
@@
 			\hline
-			CSV             & 18.931802              & 16411               \\
-			MessagePack     & 11.220957              & 15565               \\
-			CBOR            & 16.957873              & 25507               \\
-			JSON            & 21.912745              & 31180               \\
-			XML             & 21.873043              & 43699               \\
-			XBF             & 11.322245              & 14686               \\
+			CSV             & 18.93                  & 16,411              \\
+			MessagePack     & 11.22                  & 15,565              \\
+			CBOR            & 16.95                  & 25,507              \\
+			JSON            & 21.91                  & 31,180              \\
+			XML             & 21.87                  & 43,699              \\
+			XBF             & 11.32                  & 14,686              \\
 			\hline
@@
 			\hline
-			CSV             & 1823                      & 11.1084                   \\
-			MessagePack     & 977                       & 6.2769                    \\
-			CBOR            & 10919                     & 42.8079                   \\
-			JSON            & 16592                     & 53.2136                   \\
-			XML             & 29111                     & 66.6171                   \\
-			XBF             & 98                        & 0.6673                    \\
+			CSV             & 1,823                     & 11.1                      \\
+			MessagePack     & 977                       & 6.22                      \\
+			CBOR            & 10,919                    & 42.8                      \\
+			JSON            & 16,592                    & 53.2                      \\
+			XML             & 29,111                    & 66.6                      \\
+			XBF             & 98                        & 0.67                      \\
 			\hline
@@
 
-Original CSV data size recorded by the server: 17160 bytes.
+Original CSV data size recorded by the server: 17,160 bytes.
 
-Native data size recorded by the server: 14558 bytes.
+Native data size recorded by the server: 14,558 bytes.
 
 \section{Discussion}
+
+The results obtained indicate that there is still significant room for improvement in the formats used for data interchange.
+XBF performed significantly faster than all except one of the other formats tested, yet there has not been significant effort into optimizing the reference implementation's runtime performance.
+Additionally, XBF has the smallest overhead of the formats tested, indicating that it would make good choice for IOT (internet of things) systems which need to transfer data in a lower bandwidth scenario.
 

### 2023-08-11 docs/paper.tex
Update paper
Push latest paper changes

@@
 Similar to CSV (cite here?), we send a single header up front, followed by all the data.
-For a primitive type there isn't much savings here, but there are massive savings with vectors and structs.
+For a primitive type there isn't a large savings here, but there are massive savings with vectors and structs.
 With a vector the metadata is a discriminant indicating a vector is being sent, which is then followed by the metadata for its internal type.
@@
 
-The reference implementation for XBF\cite{xbf_impl} was written in Rust\cite{rust}. Rust was chosen for the following reasons:
+The reference implementation for XBF\cite{xbf_impl} is written in Rust\cite{rust}. Rust was chosen for the following reasons:
 \begin{IEEEitemize}
@@
 	\item A built-in testing system in the standard library\cite{testing}.
+	\item Errors are represented as values rather than throwing exceptions\cite{results}.
 \end{IEEEitemize}
@@
 			F64               & f64                         & double                                    \\
-			Bytes             & Vec\textless u8\textgreater & std::vector\textless u8\textgreater \\
+			Bytes             & Vec\textless u8\textgreater & std::vector\textless uint8\_t\textgreater \\
 			String            & String                      & std::string                               \\
@@
 
-Metadata for integers was an enumeration of the possible metadata discriminant values, ranging from 0 to 16, represented as unsigned 8-bit integers.
+Metadata for integers is an enumeration of the possible primitive metadata discriminant values, ranging from 0 to 16, represented as unsigned 8-bit integers.
 
@@
 
-XbfVec was implemented as native dynamic array of XbfType values, and a reference counted pointer to the metadata of its inner type.
+XbfVec is implemented as a native dynamic array of XbfType values, and a Rc (reference counted, shared pointer) to the metadata of its inner type, represented as the base XbfMetadata.
 In order to ensure XbfVec remains homogenous, it requires the metadata to be passed in at construction time.
+If all the values given to the constructor do not have the same metadata as the metadata passed in, the constructor returns an error type.
+
+The metadata is implemented with a Rc because Rust's tagged unions are not allowed to be recursive, as then they could be infinitely sized on the stack.
+A pointer type has a known sized value, and as such the tagged union can then be created on the stack by the compiler.
+The recursion comes from the base XbfMetadata type, which can contain XbfVecMetadata, which contains an XbfMetadata, and the recursion can continue.
+
+XbfVecMetadata could have been implemented with an owning pointer (known as a Box\textless T\textgreater), however that would mean that the metadata couldn't be shared among many instances of an object.
+When XbfVec is deserialized, it passes the internal metadata to each of the inner values it deserializes and constructs, which causes the reference count to increase and all the inner values to share the same heap allocated metadata.
+While this does not affect the on the wire performance, this does reduce memory usage for a user of the library.
 
 \subsection{XbfStruct}\label{struct}
+
+XbfStruct is implemented in a very similar manner to XbfVec, consisting of a Rc to the metadata of the struct, and a Box\textless [XbfType]\textgreater\space containing the fields.
+This type is known as a boxed slice.
+In Rust, a slice is a pair of a pointer to contiguous memory and a length.
+When a slice is boxed, it is allocated on the heap, and the box object is the sole owner of that heap allocation, and will de-allocate it when the box goes out of scope.
+The reason a boxed slice was chosen instead of a dynamic array was to save on memory usage.
+
+A Vec\textless T\textgreater (dynamic array) in Rust consists of a pointer to memory, a capacity of how much memory is available at that address, and a length of how many elements have been properly initialized.
+The amount of fields of a struct are not allowed to change, so a boxed slice can be used instead of a dynamic array, thereby saving the size of a pointer of memory for every XbfStruct created.
+
+The metadata of structs (XbfStructMetadata) is the most complex of the three metadata types.
+It consists of boxed string slice and an IndexMap\cite{indexmap} mapping string slices to XbfMetadata types.
+
+In Rust there is an important distinction between an owned String, and a string slice\cite{strings}.
+A string slice is similar to the previously mentioned slices, except it consists of valid UTF-8 code points.
+It is a view into a block of memory, and its size cannot be modified.
+If growable string is required, that's where the String type is used.
+It makes the same UTF-8 guarantee as a string slice, but has the same properties as a Vec where it keeps track of its capacity, will reallocate if its capacity is exceeded, similar to a dynamic array.
+As with the decision to use a slice in XbfStruct, a string slice is utilized to save on memory usage, as the metadata for a struct shouldn't need to change, thereby making a capacity field irrelevant.
+
+An IndexMap is a hash table with consistent order and fast iteration.
+Ordinarily a hash table does not keep track of the order it's fields are inserted, which is a problem when serializing the metadata, as the fields could be sent in any order.
+A dynamic array of objects consisting of the field name and type could be used, however this leads to $O(n)$ searches for a given field.
+In order to retain $O(1)$ searches, an IndexMap was used.
+It requires extra memory overhead to keep track of the insertion order of keys and values, but ensures that iteration is fast and in a consistent order.
+
+\subsection{Testing}\label{testing}
 
@@
 
-The results obtained indicate that there is still significant room for improvement in the formats used for data interchange.
-XBF performed significantly faster than all except one of the other formats tested, yet there has not been significant effort into optimizing the reference implementation's runtime performance.
-Additionally, XBF has the smallest overhead of the formats tested, indicating that it would make good choice for IOT (internet of things) systems which need to transfer data in a lower bandwidth scenario.
+XBF performed significantly faster than all except one of the other formats tested, despite minimal effort spent optimizing the reference implementation's runtime performance.
+Additionally, XBF has the smallest overhead of the formats tested, indicating that it would make a good choice for IOT (internet of things) systems which need to transfer data in a low bandwidth scenario.
+Based on the overall
 
 \section{Conclusion}
+
+The results obtained indicate that there is still significant room for improvement in the formats used for data interchange.
+The primary technique of metadata deduplication shows the overhead of sending self-describing data can be reduced greatly.
+This technique could also be applied to a human-readable format such as JSON or XML, and would provide similar benefits even if the data is transferred as text instead of binary.
+
+In the self-describing binary format space, there are still many improvements that could be made in the future to XBF. Sizes for variable length
 

### 2023-08-15 docs/paper.tex
Update paper

@@
 \usepackage{textcomp}
-\usepackage{xurl}
 \usepackage{xcolor}
 \usepackage{listings}
-\usepackage{microtype}
+\usepackage{microtype} 	% fixes word distribution in citations
+\usepackage{xurl} 			% allows citation urls to wrap around lines
+\usepackage{bytefield} 	% for shoing the memory layout of types 
+\usepackage{xpatch} 		% fixes bytefield looking wonky with IEEE formatting
 
+\xpretocmd{\bytefield}{\setlength{\lineskip}{1pt}}{}{}
+
+% weird IEEE stuff included in the template, not questioning it
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
 T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
 \begin{document}
@@
 These formats have two key features that have led to their popularity, human-readability and being self-describing.
-These features have inherent inefficiencies, which XBF (Xtensible Binary Format)\cite{xbf_spec} intends to solve through carefully chosen compromises.
+These features have inherent inefficiencies, which XBF (Xtensible Binary Format) intends to solve through carefully chosen compromises.
 
@@
 
-A self-describing format is a format in which a recipient is not required to have a schema for the incoming data, and the names and types of the data being sent can be determined from the data itself.
+A self-describing format contains metadata and does not require a schema for incoming data, and the names and types of the data being sent can be determined from the data itself.
 The following is an example of XML:
@@
 A human-readable format is one where data is encoded entirely as ASCII or UTF-8 characters, and formatted in such a way that a person could easily read or write it without the need of a machine.
-Human-readability brings with it two inefficiencies, the need to parse text data into a binary format that a computer understands, and additional overhead from whitespace characters.
+Human-readability brings with it two inefficiencies, the need to parse text data into a binary format that a computer understands, and additional overhead from the inherent amount of space it takes to encode data in ASCII.
+For example, a 32-bit number is stored in 4 bytes natively, but a variable amount of space when in ASCII.
+If the number being stored is 1, then ASCII only takes a single byte (only a single digit).
+Once more than 4 digits is required (numbers over 9999), ASCII will always be less efficient than a natively encoded number.
+In a worst case scenario, 32-bit numbers can store a number up to around 4.2 billion, which would take 10 bytes to store in plain text.
+
+\iffalse
+	\begin{table}[htbp]
+		\caption{Plain Text vs Binary}\label{binary_vs_text}
+		\begin{center}
+			\begin{tabular}{|c|c|c|}
+				\hline
+				\textbf{Value} & \textbf{Plain Text Bytes} & \textbf{Binary Bytes} \\
+				1              & 1                         & 4                     \\
+				9999           & 4                         & 4                     \\
+				10000          & 5                         & 4                     \\
+				\hline
+			\end{tabular}
+		\end{center}
+	\end{table}
+\fi
 
@@
 
-Similar to CSV (cite here?), we send a single header up front, followed by all the data.
-For a primitive type there isn't a large savings here, but there are massive savings with vectors and structs.
-With a vector the metadata is a discriminant indicating a vector is being sent, which is then followed by the metadata for its internal type.
-This process continues recursively if the inner type (or any of its inner types) contains a vector or struct, ensuring that all metadata necessary to represent the data coming afterward is received up front.
-The data is then sent in a single continuous block, something after this...
+\subsection{Primitive Types}
+
+There are a total of 17  primitive types:
+
+\begin{IEEEitemize}
+	\item Boolean
+	\item U8, U16, U32, U64, U128, U256
+	\item I8, I16, I32, I64, I128, I256
+	\item F32, F64
+	\item Bytes
+	\item String (UTF-8)
+\end{IEEEitemize}
+
+All integers should be sent in little endian format, with the least significant bit first.
+For signed integers, they should be represented in two's complement format.
+These two formats are chosen because they are the most common way integers are interacted with in modern x86 and ARM processors, and should minimize the amount of conversion required when sending and receiving data.
+Similarly, floating point numbers should be sent as 32 or 64 bit IEEE 754 floating point numbers.
+
+Strings should be sent as a sequence of bytes that correspond to UTF-8 code points.
+They should first send their length as an unsigned 64-bit integer (in little endian format), followed by the corresponding number of bytes contained within the string.
+Bytes have the same specification as strings, but with the exception that they do not have to be a valid sequence of UTF-8 code points.
+Lengths were chosen to be 64-bits to simplify the format, but at the cost of efficiency when sending many small sequences.
+
+\subsection{Primitive Metadata}
+
+Primitive metadata should be sent as a single byte discriminant value.
+After receiving one of these discriminant values, the client should read the following byte(s) and interpret them as the type given by the discriminant.
+
+\begin{table}[htbp]
+	\caption{Primitive Metadata Discriminants}\label{discriminants}
+	\begin{center}
+		\begin{tabular}{|c|c|}
+			\hline
+			\textbf{Type} & \textbf{Discriminant} \\
+			\hline
+			Boolean       & 0                     \\
+			U8            & 1                     \\
+			U16           & 2                     \\
+			U32           & 3                     \\
+			U64           & 4                     \\
+			U128          & 5                     \\
+			U256          & 6                     \\
+			I8            & 7                     \\
+			I16           & 8                     \\
+			I32           & 9                     \\
+			I64           & 10                    \\
+			I128          & 11                    \\
+			I256          & 12                    \\
+			F32           & 13                    \\
+			F64           & 14                    \\
+			Bytes         & 15                    \\
+			String        & 16                    \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+Strings should always be the final value in the list. The value given to strings
+is used by Vectors and Structs to determine what their discriminant value should
+be.
+
+\subsection{Primitive Examples}
+
+In the following examples, each type will first be shown with its discriminant value followed by its actual value, with the byte representation underneath them.
+
+\subsubsection{16-Bit Unsigned Integer}
+
+\begin{center}
+	\begin{bytefield}{24}
+		\bitbox{8}{Disc} & \bitbox{16}{1024} \\
+		\bitbox{8}{2} & \bitboxes{8}{04} \\
+	\end{bytefield}
+\end{center}
+
+\subsubsection{String}
+
+\newcommand{\rot}[1]{\rotatebox{90}{#1}}
+
+\begin{center}
+	\begin{bytefield}[bitheight=\widthof{"Disc"}]{28}
+		\bitbox{2}{\rot{Disc}} & \bitbox{16}{length} & \bitboxes{2}{hello} \\
+		\bitboxes{2}{{16} {2} {0} {0} {0} {0} {0} {0} {0} {\rot{104}} {\rot{101}} {\rot{108}} {\rot{108}} {\rot{111}}}
+	\end{bytefield}
+\end{center}
+
+\subsection{Vector Type}
+
+Vectors are a homogenous list of values that has a known, variable length.
+Vectors should first include their length as an unsigned 64-bit integer (the same as Strings and Bytes), followed by the corresponding number of elements.
+The type contained within a Vector is not sent to the client.
+That information is carried in the metadata.
+
+\subsection{Vector Metadata}
+
+Similarly to primitives, a single byte discriminant value signifying a Vector is incoming is sent.
+This discriminant value should be 1 greater than that of the discriminant value for Strings.
+Following this, metadata information for the internal type contained within the Vector will be sent.
+This process may continue recursively with nested types of Vectors and Structs.
+The length of a vector or the data contained within the vector must not be sent.
+
+\subsection{Vector Example}
+
+\textit{Vector of 16-Bit Unsigned Integers}
+
+DV = Discriminant of a Vector
+
+D16 = Discriminant of a 16-Bit Unsigned Integer
+
+\begin{center}
+	\begin{bytefield}{26}
+		\bitbox{2}{\rot{DV}} & \bitbox{2}{\rot{D16}} & \bitbox{16}{length} & \bitboxes{4}{{42} {1024}} \\
+		\bitboxes{2}{{17} {2} {2} {0} {0} {0} {0} {0} {0} {0} {42} {0} {0} {4}}
+	\end{bytefield}
+\end{center}
+
+\subsection{Struct Type}
+
+A Struct is an aggregate type containing a name as well as named fields.
+A struct may not contain duplicate field names.
+Should a Struct be sent like this anyway, it should be considered malformed and not be constructed on the receiving end.
+Fields of a Struct are sent in sequence in the order they are listed in the Struct's metadata.
+When a struct is serialized it should not send any sort of name information (such as its name or field names), how many fields it has, nor should it send any type information about its fields.
+That information is carried in the metadata.
+
+\subsection{Struct Metadata}
+
+A discriminant value should first be sent, similarly to primitives (following the same size requirement). This discriminant value should be 1 greater than that of the discriminant value for Vectors.
+
+Following this, the name of the Struct should be sent, using the same format as primitive strings are sent (unsigned 64-bit length and then the bytes).
+Next, send the number of fields contained within the Struct as a U64, the same as all other lengths.
+Finally, the fields of the Struct should be sent, first the name of the field as a String, then immediately after the metadata for the type of the field.
+This process may continue recursively with nested types of Structs or Vectors.
+These name and type pairs will be sent until there are no more fields left in the Struct.
+
+\subsection{Struct Example}
+
+\begin{lstlisting}[language=C]
+struct MyStruct {
+	x: i8,
+	y: i8
+}
+
+let my_struct = MyStruct {
+	x: 42,
+	y: -5,
+};
+\end{lstlisting}
 
@@
 The reference implementation for XBF\cite{xbf_impl} is written in Rust\cite{rust}. Rust was chosen for the following reasons:
+
 \begin{IEEEitemize}
@@
 In order to perform a downcast on either of these top level types, Rust's structural pattern matching must be used.
+
+\begin{table}[htbp]
+	\begin{center}
+		\begin{tabular}{|c|c|c|}
+			\hline
+			\multicolumn{3}{|c|}{XbfType}     \\
+			\hline
+			XbfPrimitive & XbfVec & XbfStruct \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+\begin{table}[htbp]
+	\begin{center}
+		\begin{tabular}{|c|c|c|}
+			\hline
+			\multicolumn{3}{|c|}{XbfMetadata}                         \\
+			\hline
+			XbfPrimitiveMetadata & XbfVecMetadata & XbfStructMetadata \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
 
@@
 
+\begin{center}
+	\begin{bytefield}{32}
+		\wordbox{1}{XbfVec} \\
+		\bitbox{8}{Metadata} & \bitbox{24}{Vec\textless XbfType\textgreater} \\
+		\bitbox{8}{Rc} & \bitbox{8}{*XbfType} & \bitbox{8}{Length} & \bitbox{8}{Capacity}
+	\end{bytefield}
+\end{center}
+
 The metadata is implemented with a Rc because Rust's tagged unions are not allowed to be recursive, as then they could be infinitely sized on the stack.
@@
 
+\begin{center}
+	\begin{bytefield}{32}
+		\wordbox{1}{XbfStruct} \\
+		\bitbox{8}{Metadata} & \bitbox{24}{Box\textless [XbfType]\textgreater} \\
+		\bitbox{8}{Rc} & \bitbox{12}{*XbfType} & \bitbox{12}{Length}
+	\end{bytefield}
+\end{center}
+
 A Vec\textless T\textgreater (dynamic array) in Rust consists of a pointer to memory, a capacity of how much memory is available at that address, and a length of how many elements have been properly initialized.
@@
 If growable string is required, that's where the String type is used.
-It makes the same UTF-8 guarantee as a string slice, but has the same properties as a Vec where it keeps track of its capacity, will reallocate if its capacity is exceeded, similar to a dynamic array.
+It makes the same UTF-8 guarantee as a string slice, but has the same properties as a Vec where it keeps track of its capacity, and will reallocate if its capacity is exceeded, similar to a dynamic array.
 As with the decision to use a slice in XbfStruct, a string slice is utilized to save on memory usage, as the metadata for a struct shouldn't need to change, thereby making a capacity field irrelevant.
+
+\begin{center}
+	\begin{bytefield}{24}
+		\wordbox{1}{String} \\
+		\bitbox{8}{*u8} & \bitbox{8}{Length} & \bitbox{8}{Capacity} \\
+	\end{bytefield}
+
+	\begin{bytefield}{16}
+		\wordbox{1}{\&str (string slice)} \\
+		\bitbox{8}{*u8} & \bitbox{8}{Length}
+	\end{bytefield}
+\end{center}
 
@@
 
+The first result to address is the difference between the number of bytes in the original CSV file and the number of bytes received by the client when asking for CSV data.
+Upon examination, it was found that the original CSV always contained 8 total digits of precision, whereas when it was serialized only the number of digits required to represent a number was sent.
+As an example, in the original data the number 83.5 would be represented as 83.500000, but when serialized would omit the trailing zeroes, as they were unnecessary.
+
 XBF performed significantly faster than all except one of the other formats tested, despite minimal effort spent optimizing the reference implementation's runtime performance.
+This can likely be attributed working entirely with binary data, which avoids needing to do complex and expensive conversions between plain text and native binary formats, as well as requires sending less data in general.
+TODO: more here
+
 Additionally, XBF has the smallest overhead of the formats tested, indicating that it would make a good choice for IOT (internet of things) systems which need to transfer data in a low bandwidth scenario.
-Based on the overall
+Based on the overall TODO: more here
 
@@
 
-In the self-describing binary format space, there are still many improvements that could be made in the future to XBF. Sizes for variable length
-
 \bibliographystyle{IEEEtran}

### 2023-08-15 docs/paper.tex
Update paper

@@
 \begin{abstract}
-	Do later
+	The most common data interchange formats used on the web are self-describing, and include metadata with every entry sent.
+Additionally, they are also encoded in plain text.
+Both of these traits are inefficient, and data transfer size and speed can be massively improved with a new format.
+This paper seeks to define and implement a specification for a new data interchange format that remains self-describing, but heavily deduplicates metadata and transfers data in binary.
+This allows it to be as fast, or faster, than other data interchange formats, even those that utilize binary encodings, while requiring overall less bytes sent on the wire.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-	Data format, XML, JSON, Binary data
+	JSON, XML, Binary Encoding, Data Interchange
 \end{IEEEkeywords}
@@
 In this particular example, 14 characters of the 59 total characters actually contain the desired data, the rest are just describing the metadata.
-The best thing a format can do is minimize this overhead as much as possible, it cannot be eliminated while still being self-describing.
 
@@
 A human-readable format is one where data is encoded entirely as ASCII or UTF-8 characters, and formatted in such a way that a person could easily read or write it without the need of a machine.
-Human-readability brings with it two inefficiencies, the need to parse text data into a binary format that a computer understands, and additional overhead from the inherent amount of space it takes to encode data in ASCII.
+Human-readability brings with it two inefficiencies, the need to parse text data into a binary format that a computer understands, and additional overhead from the inherent amount of space it takes to encode data in text.
 For example, a 32-bit number is stored in 4 bytes natively, but a variable amount of space when in ASCII.
@@
 In a worst case scenario, 32-bit numbers can store a number up to around 4.2 billion, which would take 10 bytes to store in plain text.
-
-\iffalse
-	\begin{table}[htbp]
-		\caption{Plain Text vs Binary}\label{binary_vs_text}
-		\begin{center}
-			\begin{tabular}{|c|c|c|}
-				\hline
-				\textbf{Value} & \textbf{Plain Text Bytes} & \textbf{Binary Bytes} \\
-				1              & 1                         & 4                     \\
-				9999           & 4                         & 4                     \\
-				10000          & 5                         & 4                     \\
-				\hline
-			\end{tabular}
-		\end{center}
-	\end{table}
-\fi
 
@@
 Following this, the name of the Struct should be sent, using the same format as primitive strings are sent (unsigned 64-bit length and then the bytes).
-Next, send the number of fields contained within the Struct as a U64, the same as all other lengths.
+Next, send the number of fields contained within the Struct as an unsigned 16-bit integer.
+A 16-bit integer is used because it's very unlikely for a struct of more than 65 thousand fields to be necessary.
 Finally, the fields of the Struct should be sent, first the name of the field as a String, then immediately after the metadata for the type of the field.
@@
 
+The following pseudocode will be used as an example for the description of a struct:
+
 \begin{lstlisting}[language=C]
-struct MyStruct {
-	x: i8,
-	y: i8
+struct Point {
+	first: i8,
+	last: i8
 }
 
-let my_struct = MyStruct {
-	x: 42,
-	y: -5,
+let my_point = Point {
+	first: 42,
+	last: -5,
 };
 \end{lstlisting}
+
+The struct has a name "Point", and two fields, named "first" and "last", both of type i8. The struct's metadata should be sent as follows:
+
+\begin{center}
+	\begin{bytefield}{32}
+		\bitbox{2}{\rot{DS}} & \bitbox{16}{struct name length} & \bitboxes{2}{Point} & \bitbox{4}{fields} \\
+		\bitboxes{2}{{18} {5} {0} {0} {0} {0} {0} {0} {0} {80} {\rot{111}} {\rot{105}} {\rot{110}} {\rot{116}} {2} {0}} \\
+		\bitbox{16}{first name length} & \bitboxes{2}{first} & \bitbox{2}{\rot{DX}} \\
+		\bitboxes{2}{{5} {0} {0} {0} {0} {0} {0} {0} {\rot{102}} {\rot{105}} {\rot{114}} {\rot{115}} {\rot{116}} {7}} \\
+		\bitbox{16}{last name length} & \bitboxes{2}{last} & \bitbox{2}{\rot{DY}} \\
+		\bitboxes{2}{{4} {0} {0} {0} {0} {0} {0} {0} {\rot{108}} {\rot{97}} {\rot{115}} {\rot{116}} {7}}
+	\end{bytefield}
+\end{center}
 
@@
 
-\section{Discussion}
+\section{Conclusions}
+
+\subsection{CSV Size Discrepancy}
 
@@
 As an example, in the original data the number 83.5 would be represented as 83.500000, but when serialized would omit the trailing zeroes, as they were unnecessary.
+This difference only serves to increase the performance of CSV in the test, so was left as is.
 
-XBF performed significantly faster than all except one of the other formats tested, despite minimal effort spent optimizing the reference implementation's runtime performance.
-This can likely be attributed working entirely with binary data, which avoids needing to do complex and expensive conversions between plain text and native binary formats, as well as requires sending less data in general.
-TODO: more here
+\subsection{Time Performance}
 
-Additionally, XBF has the smallest overhead of the formats tested, indicating that it would make a good choice for IOT (internet of things) systems which need to transfer data in a low bandwidth scenario.
-Based on the overall TODO: more here
+XBF performed significantly faster than all other formats except for MessagePack, in which both formats performed in about the same amount of time.
+This is likely due to both formats working with binary data, which avoids needing to do complex and expensive conversions between plain text and native binary formats, as well as requires sending less data in general.
 
-\section{Conclusion}
+The XBF reference implementation prioritized correctness, memory usage, and performance, in that order.
+Most optimization time was put into reducing the number of heap allocations made as they are relatively expensive.
+Given more time the program could be profiled to discover where the most execution time is spent, and optimizations made based on that to improve runtime performance even further.
 
-The results obtained indicate that there is still significant room for improvement in the formats used for data interchange.
-The primary technique of metadata deduplication shows the overhead of sending self-describing data can be reduced greatly.
-This technique could also be applied to a human-readable format such as JSON or XML, and would provide similar benefits even if the data is transferred as text instead of binary.
+\subsection{Size Performance}
+
+Of all the formats tested, XBF was able to send the least amount of bytes to express the same amount of data.
+This is despite some of the optimizations XBF is missing compared to the next most efficient format, MessagePack.
+MessagePack allows for the conversion of numbers to as small of a size as possible before being sent over the wire.
+For example, if an unsigned 64-bit number with a value of 1024 was to be sent, it would be downgraded to an unsigned 16-bit number as no information is lost from doing so.
+MessagePack also has different string and array types with different integer length types, allowing fewer bytes to be sent when these variable length types don't have many elements.
+As an example, an array with less than 65,536 elements would be sent as an array 16, instead of an array 32.
+Both of these optimizations weren't implemented in the XBF specification or reference implementation for the sake of simplicity, but could be added in a future revision.
+
+XBF requiring the least amount of bytes, as well as having the smallest overhead, indicates that the primary technique of metadata deduplication worked as intended.
+This could be implemented in other formats as well, including the human-readable formats such as JSON and XML.
+
+In its current state, XBF would likely make a good candidate for use in low bandwidth scenarios, such as IoT (Internet of Things) systems.
+It could also have applications in extremely high volume situations, where even a difference of a few kilobytes in the amount of data sent could add up to gigabytes saved very quickly.
 

### 2023-08-15 docs/paper.tex
Update paper

@@
 
-The struct has a name "Point", and two fields, named "first" and "last", both of type i8. The struct's metadata should be sent as follows:
+The struct has a name Point, and two fields, named first and last, both of type i8.
+The goal is to serialize the variable \texttt{my\_point}, which is an instance of Point with the values 42 and -5 for the members first and last respectively.
+The struct's metadata should be sent as follows:
 
@@
 		\bitboxes{2}{{4} {0} {0} {0} {0} {0} {0} {0} {\rot{108}} {\rot{97}} {\rot{115}} {\rot{116}} {7}}
+	\end{bytefield}
+\end{center}
+
+The values inside the struct should then be sent in the same order as they are listed in the struct's metadata:
+
+\begin{center}
+	\begin{bytefield}{12}
+		\bitbox{6}{first} & \bitbox{6}{last} \\
+		\bitbox{6}{42} & \bitbox{6}{251} \\
 	\end{bytefield}
@@
 The metadata of structs (XbfStructMetadata) is the most complex of the three metadata types.
-It consists of boxed string slice and an IndexMap\cite{indexmap} mapping string slices to XbfMetadata types.
+It consists of boxed string slice and an IndexMap\cite{indexmap} mapping boxed string slices to XbfMetadata types.
 
@@
 It makes the same UTF-8 guarantee as a string slice, but has the same properties as a Vec where it keeps track of its capacity, and will reallocate if its capacity is exceeded, similar to a dynamic array.
-As with the decision to use a slice in XbfStruct, a string slice is utilized to save on memory usage, as the metadata for a struct shouldn't need to change, thereby making a capacity field irrelevant.
+As with the decision to use a boxed slice in XbfStruct, a boxed string slice is utilized to save on memory usage, as the metadata for a struct shouldn't need to change, thereby making a capacity field irrelevant.
 
@@
 
-\subsection{Testing}\label{testing}
+\begin{center}
+	\begin{bytefield}{36}
+		\wordbox{1}{XbfStructMetadata} \\
+		\bitbox{10}{Box\textless str\textgreater} & \bitbox{26}{IndexMap\textless Box\textless str\textgreater, XbfMetadata\textgreater} \\
+		\bitbox{5}{*u8} & \bitbox{5}{Length} & \bitbox{26}{Fields omitted for brevity}
+	\end{bytefield}
+\end{center}
+
+\subsection{Implementation Testing}\label{testing}
+
+The reference implementation was developed in a test driven manner, taking advantage of the built-in testing facilities of Rust.
+Individual unit tests would first be written with the desired behavior for an object or function, and then code would be written to satisfy the test.
+This approach leads to more reliable software.
+When adding a new feature, if anything from old features is broken tests will fail, indicating something has gone wrong.
+Unit tests also caught a number of implementation bugs which were noticed immediately because of the immediate feedback.
+
+In addition to the unit tests, a number of integration tests, tests designed to test the public API (Application Programming Interface), were implemented to ensure the experience of using the library was as expected.
+The combination of these approaches has led to 100 percent code coverage, such that every line of source code of the library has been executed at least once by a test.
 

### 2023-08-16 docs/paper/paper.tex
Update paper, move stuff to separate folders

--- 2023-08-15 docs/paper.tex
@@
 
-On the web, data is largely interchanged in one of two formats, XML (Extensible Markup Language)\cite{xml_spec} or JSON (JavaScript Object Notation)\cite{json_spec}.
-These formats have two key features that have led to their popularity, human-readability and being self-describing.
+On the web, data is largely interchanged in one of two formats, Extensible Markup Language (XML)\cite{xml_spec} or  JavaScript Object Notation (JSON)\cite{json_spec}.
+These formats have two key features that have led to their popularity: self-description and human-readability.
 These features have inherent inefficiencies, which XBF (Xtensible Binary Format) intends to solve through carefully chosen compromises.
@@
 
-An additional optimization that could be made (but isn't with either of these formats) is deduplication of the metadata.
+% An additional optimization that could be made (but isn't with either of these formats) is deduplication of the metadata.
 Both JSON and XML include the metadata for every element sent, even if what is being sent is a homogenous list of elements that all share the same structure.
+%
 
-\subsection{Human Readability}
+\subsection{Human-Readability}
 
@@
 
-\subsection{Enhancements Made}
+\subsection{Similar Formats}
 
-With the inefficiencies of these two formats in mind, XBF seeks to improve on them through a few core design decisions:
+There are other self-describing formats that improve on JSON and XML in various ways that will now be examined.
+
+\subsubsection{Comma Separated Values (CSV)}
+
+CSV\cite{csv_spec} is a human-readable format ideal for sending mass amounts of data.
+CSV avoids sending metadata repeatedly by having a header at the top of the file which describes the names of each of the columns in the rows that follow.
+These rows must always contain the same number of elements as the number of columns defined in the header, which makes CSV less flexible compared to JSON and XML, which can have objects that contain objects or lists of objects.
+
+\subsubsection{Concise Binary Object Format (CBOR)}
+
+CBOR\cite{cbor_spec} is a binary format which seeks to have interoperability with JSON, a fairly small message size, and tiny required code size for an encoder and decoder.
+Additionally, the format seeks to be future-proof and extensible such that older decoders can still decode new extended data.
+
+\subsubsection{MessagePack}
+
+MessagePack\cite{msgpack_spec} is a binary format inspired by JSON but with its own type system and encoding and decoding format.
+All data is represented in a binary format, big-endian where necessary, and supports custom extension types that can be defined by applications.
+
+\subsection{Shortcomings}
+
+Each of the formats examined has its own set of shortcomings that can be improved on.
+CSV has the right idea of sending as little metadata as possible, but has a limited type system as a result.
+JSON and XML are human-readable, and thus inefficient by design.
+CBOR prioritizes implementation code size more than a small message size, when it should be possible to achieve both.
+MessagePack doesn't deduplicate metadata where possible, and sends type information with every value sent.
+Additionally, MessagePack utilizes big-endian formatting, while most modern processors use little-endian, requiring a conversion to take place.
+
+Given these shortcomings, there are a number of ways they can be improved on:
 
 \begin{IEEEitemize}
-	\item Send data and metadata in a binary format (no longer human-readable).
-	\item Remain self-describing, but have the option to not include metadata if it's already known, and deduplicate metadata whenever possible.
-	\item Prioritize simplicity in the type system.
+	\item When possible, send metadata only a single time (mainly for homogenous lists).
+	\item Allow sending metadata to be optional, such as if a client already has it.
+	\item Utilize little-endian to reduce conversions.
+	\item Prioritize a simple type system.
 \end{IEEEitemize}
@@
 
-All integers should be sent in little endian format, with the least significant bit first.
-For signed integers, they should be represented in two's complement format.
-These two formats are chosen because they are the most common way integers are interacted with in modern x86 and ARM processors, and should minimize the amount of conversion required when sending and receiving data.
-Similarly, floating point numbers should be sent as 32 or 64 bit IEEE 754 floating point numbers.
+All integers will be sent in little endian format, with the least significant bit first.
+For signed integers, they will be represented in two's complement format.
+These two formats are chosen because they are the most common way integers are interacted with in modern x86 and ARM processors, and will minimize the amount of conversion required when sending and receiving data.
+Similarly, floating point numbers will be sent as 32 or 64 bit IEEE 754 floating point numbers.
 
-Strings should be sent as a sequence of bytes that correspond to UTF-8 code points.
-They should first send their length as an unsigned 64-bit integer (in little endian format), followed by the corresponding number of bytes contained within the string.
+Strings will be sent as a sequence of bytes that correspond to UTF-8 code points.
+They will first send their length as an unsigned 64-bit integer (in little endian format), followed by the corresponding number of bytes contained within the string.
 Bytes have the same specification as strings, but with the exception that they do not have to be a valid sequence of UTF-8 code points.
@@
 
-Primitive metadata should be sent as a single byte discriminant value.
-After receiving one of these discriminant values, the client should read the following byte(s) and interpret them as the type given by the discriminant.
+Primitive metadata will be sent as a single byte discriminant value.
+After receiving one of these discriminant values, the client will read the following byte(s) and interpret them as the type given by the discriminant.
+The discriminant for each primitive type is show in Table \ref{discriminants}
 
@@
 The struct's metadata should be sent as follows:
+
+DS = Discriminant of a Struct
+
+DX = Discriminant of field "first"
+
+DY = Discriminant of field "last"
 
@@
 \begin{IEEEitemize}
-	\item CSV (Comma Separated Values)\cite{csv_spec}\cite{csv_parser}
-	\item MessagePack\cite{msgpack_spec}\cite{msgpack_parser}
-	\item CBOR (Concise Binary Object Representation)\cite{cbor_spec}\cite{cbor_parser}
+	\item CSV\cite{csv_parser}
+	\item MessagePack\cite{msgpack_parser}
+	\item CBOR\cite{cbor_parser}
 	\item JSON\cite{json_parser}

### 2023-08-16 docs/presentation/presentation.tex
Update paper, move stuff to separate folders

\documentclass{beamer}
\usepackage{amsfonts,amsmath,oldgerm}
\usetheme{sintef}

\newcommand{\testcolor}[1]{\colorbox{#1}{\textcolor{#1}{test}}~\texttt{#1}}

\usefonttheme[onlymath]{serif}

\titlebackground*{assets/background}

\newcommand{\hrefcol}[2]{\textcolor{cyan}{\href{#1}{#2}}}

\title{Implementing XBF: An Efficient Self-Describing Binary Format}
% \subtitle{Using \LaTeX\ to prepare slides}
% \course{Master's Degree in Computer Science}
\author{\href{mailto:dkrautha@pm.me}{David Krauthamer}}
% \IDnumber{1234567}

\begin{document}
\maketitle

\begin{frame}

	This template is based on \hrefcol{https://www.overleaf.com/latex/templates/sapienza-presentation/yyksmrskffhk}{Sapienza Presentation} from \hrefcol{mailto:andrea@gasparini.cloud}{Andrea Gasparini}

	\vspace{\baselineskip}

	The above template is a based on \hrefcol{https://www.overleaf.com/latex/templates/sintef-presentation/jhbhdffczpnx}{SINTEF Presentation} from \hrefcol{mailto:federico.zenith@sintef.no}{Federico Zenith} and its derivation \hrefcol{https://github.com/TOB-KNPOB/Beamer-LaTeX-Themes}{Beamer-LaTeX-Themes} from Liu Qilong

	\vspace{\baselineskip}

	In the following you find a brief introduction on how to use \LaTeX\ and the beamer package to prepare slides, based on the one written by \hrefcol{mailto:federico.zenith@sintef.no}{Federico Zenith} for \hrefcol{https://www.overleaf.com/latex/templates/sintef-presentation/jhbhdffczpnx}{SINTEF Presentation}

	% This template is released under \hrefcol{https://creativecommons.org/licenses/by-nc/4.0/legalcode}{Creative Commons CC BY 4.0} license

\end{frame}

\section{Introduction}

\begin{frame}{Beamer for SINTEF slides}
	\begin{itemize}
		\item We assume you can use \LaTeX; if you cannot,
		      \hrefcol{http://en.wikibooks.org/wiki/LaTeX/}{you can learn it here}
		\item Beamer is one of the most popular and powerful document
		      classes for presentations in \LaTeX
		\item Beamer has also a detailed
		      \hrefcol{http://www.ctan.org/tex-archive/macros/latex/contrib/beamer/doc/beameruserguide.pdf}{user
			      manual}
		\item Here we will present only the most basic features to get you up to speed
	\end{itemize}
\end{frame}

\begin{frame}{Beamer vs. PowerPoint}
	Compared to PowerPoint, using \LaTeX\ is better because:
	\begin{itemize}
		\item It is not What-You-See-Is-What-You-Get, but
		      What-You-\emph{Mean}-Is-What-You-Get:\\
		      you write the content, the computer does the typesetting
		\item Produces a \texttt{pdf}: no problems with fonts, formulas,
		      program versions
		\item Easier to keep consistent style, fonts, highlighting, etc.
		\item Math typesetting in \TeX\ is the best:
		      \begin{equation*}
			      \mathrm{i}\,\hslash\frac{\partial}{\partial t} \Psi(\mathbf{r},t) =
			      -\frac{\hslash^2}{2\,m}\nabla^2\Psi(\mathbf{r},t)
			      + V(\mathbf{r})\Psi(\mathbf{r},t)
		      \end{equation*}

	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Getting Started}
	\framesubtitle{Selecting the SINTEF Theme}
	To start working with \texttt{sintefbeamer}, start a \LaTeX\ document with the
	preamble:
	\begin{block}{Minimum SINTEF Beamer Document}
		\verb|\documentclass{beamer}|\\
		\verb|\usetheme{sintef}|\\
		\verb|\begin{document}|\\
		\verb|\begin{frame}{Hello, world!}|\\
		\verb|\end{frame}|\\
		\verb|\end{document}|\\
	\end{block}
\end{frame}

\begin{frame}[fragile]{Title page}
	To set a typical title page, you call some commands in the preamble:
	\begin{block}{The Commands for the Title Page}
		\begin{verbatim}
\title{Sample Title}
\subtitle{Sample subtitle}
\author{First Author, Second Author}
\date{\today} % Can also be (ab)used for conference name &c.
\end{verbatim}
	\end{block}
	You can then write out the title page with \verb|\maketitle|.

	To set a \textbf{background image} use the \verb|\titlebackground| command
	before \verb|\maketitle|; its only argument is the name (or path) of a graphic
	file.

	If you use the \textbf{starred version} \verb|\titlebackground*|, the image
	will be clipped to a split view on the right side of the title slide.

\end{frame}

\begin{frame}[fragile]{Writing a Simple Slide}
	\framesubtitle{It's really easy!}
	\begin{itemize}[<+->]
		\item A typical slide has bulleted lists
		\item These can be uncovered in sequence
	\end{itemize}
	\begin{block}{Code for a Page with an Itemised List}<+->
		\begin{verbatim}
\begin{frame}{Writing a Simple Slide}
\framesubtitle{It's really easy!}
\begin{itemize}[<+->]
\item A typical slide has bulleted lists
\item These can be uncovered in sequence
\end{itemize}\end{frame}
\end{verbatim}
	\end{block}
\end{frame}

\section{Personalization}

\footlinecolor{stevensmediumgold}
\begin{frame}[fragile]{Changing Slide Style}
	\begin{itemize}
		\item You can select the white or \textit{maincolor} \textbf{slide style} \emph{in the
			      preamble} with \verb|\themecolor{white}| (default) or \verb|\themecolor{main}|
		      \begin{itemize}
			      \item You should \emph{not} change these within the document: Beamer does
			            not like it
			      \item If you \emph{really} must, you may have to add
			            \verb|\usebeamercolor[fg]{normal text}| in the slide
		      \end{itemize}
		\item You can change the \textbf{footline colour} with
		      \verb|\footlinecolor{color}|
		      \begin{itemize}
			      \item Place the command \emph{before} a new \verb|frame|
			      \item There are four ``official'' colors:
			            \testcolor{maincolor}, \testcolor{sintefyellow},
			            \testcolor{sintefgreen}, \testcolor{sintefdarkgreen}
			      \item Default is no footline; you can restore it with
			            \verb|\footlinecolor{}|
			      \item Others may work, but no guarantees!
			      \item Should \emph{not} be used with the \verb|maincolor| theme!
		      \end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Blocks}
	\begin{columns}
		\begin{column}{0.3\textwidth}
			\begin{block}{Standard Blocks}
				These have a color coordinated with the footline (and grey in the blue theme)
				\begin{verbatim}
\begin{block}{title}
content...
\end{block}
\end{verbatim}
			\end{block}
		\end{column}
		\begin{column}{0.7\textwidth}
			\begin{colorblock}[black]{stevenslightorange}{Colour Blocks}
				Similar to the ones on the left, but you pick the colour. Text will be white by
				default, but you may set it with an optional argument.
				\small
				\begin{verbatim}
\begin{colorblock}[black]{sinteflightgreen}{title}
content...
\end{colorblock}
\end{verbatim}
			\end{colorblock}
			The ``official'' colours of colour blocks are: \testcolor{sinteflilla},
			\testcolor{maincolor}, \testcolor{sintefdarkgreen}, and
			\testcolor{sintefyellow}.
		\end{column}
	\end{columns}
\end{frame}

\footlinecolor{}
\begin{frame}[fragile]{Using Colours}
	\begin{itemize}[<alert@2>]
		\item You can use colours with the
		      \verb|\textcolor{<color name>}{text}| command
		\item The colours are defined in the \texttt{sintefcolor} package:
		      \begin{itemize}
			      \item Primary colours: \testcolor{maincolor} and its sidekick
			            \testcolor{stevensgray}
			      \item Secondary Colors are: \testcolor{stevensdarkgray}, \testcolor{stevensdarkblue},\testcolor{stevensmediumblue}, \testcolor{stevensmediumorange}, \testcolor{stevensmediumgold}
			      \item Light colors are:
			            \testcolor{stevenslightgray},\testcolor{stevenslightblue}, \testcolor{stevenslightorange}, \testcolor{stevenslightgold}
			      \item Other colors are \testcolor{sinteflightgreen},
			            \testcolor{sintefgreen}, \testcolor{sintefdarkgreen}
			      \item Additional colours: \testcolor{sintefyellow}, \testcolor{sintefred},
			            \testcolor{sinteflilla}
			            \begin{itemize}
				            \item These may be shaded---see the \verb|sintefcolor| documentation or
				                  the \hrefcol{https://sintef.sharepoint.com/sites/stottetjenester/%
					                  kommunikasjon/grafisk-profil-new/Sider/default.aspx}{SINTEF profile
					                  manual}
			            \end{itemize}
		      \end{itemize}
		\item Do \emph{not} abuse colours: \verb|\emph{}| is usually enough
		\item Use \verb|\alert{}| to bring the \alert<2->{focus} somewhere
		\item<2- | alert@2> If you highlight too much, you don't highlight at all!
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Adding images}
	\begin{columns}
		\begin{column}{0.7\textwidth}
			Adding images works like in normal \LaTeX:
			\begin{block}{Code for Adding Images}
				\begin{verbatim}
\usepackage{graphicx}
% ...
\includegraphics[width=\textwidth]
{assets/logo_RGB}
\end{verbatim}
			\end{block}
		\end{column}
		\begin{column}{0.3\textwidth}
			\includegraphics[width=\textwidth]
			{assets/logo_RGB}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}[fragile]{Splitting in Columns}
	Splitting the page is easy and common;
	typically, one side has a picture and the other text:
	\begin{columns}
		\begin{column}{0.6\textwidth}
			This is the first column
		\end{column}
		\begin{column}{0.3\textwidth}
			And this the second
		\end{column}
	\end{columns}
	\begin{block}{Column Code}
		\begin{verbatim}
\begin{columns}
\begin{column}{0.6\textwidth}
  This is the first column
\end{column}
\begin{column}{0.3\textwidth}
  And this the second
\end{column}
% There could be more!
\end{columns}
\end{verbatim}
	\end{block}
\end{frame}

\begin{chapter}[assets/background_negative]{}{Special Slides}
 \begin{itemize}
	 \item Chapter slides
	 \item Side-picture slides
 \end{itemize}
\end{chapter}

\footlinecolor{sintefred}
\begin{frame}[fragile]{Chapter slides}
	\begin{itemize}
		\item Similar to \verb|frame|s, but with a few more options
		\item Opened with \verb|\begin{chapter}[<image>]{<color>}{<title>}|
		\item Image is optional, colour and title are mandatory
		\item There are seven ``official'' colours: \testcolor{maincolor},
		      \testcolor{sintefdarkgreen}, \testcolor{sintefgreen},
		      \testcolor{sinteflightgreen}, \testcolor{sintefred}, \testcolor{sintefyellow},
		      \testcolor{sinteflilla}.
		      \begin{itemize}
			      \item Strangely enough, these are \emph{more} than the official colours
			            for the footline.
			      \item It may still be a nice touch to change the footline of following
			            slides to the same color of a chapter slide. Your choice.
		      \end{itemize}
		\item Otherwise, \verb|chapter| behaves just like \verb|frame|.
	\end{itemize}
\end{frame}

\begin{sidepic}{assets/background_alternative}{Side-Picture Slides}
	\begin{itemize}
		\item Opened with \texttt{$\backslash$begin\{sidepic\}\{<image>\}\{<title>\}}
		\item Otherwise, \texttt{sidepic} works just like \texttt{frame}
	\end{itemize}
\end{sidepic}

\footlinecolor{}
\begin{frame}
	\frametitle{Fonts}
	\begin{itemize}
		\item The paramount task of fonts is being readable
		\item There are good ones...
		      \begin{itemize}
			      \item {\textrm{Use serif fonts only with high-definition projectors}}
			      \item {\textsf{Use sans-serif fonts otherwise (or if you simply prefer
				            them)}}
		      \end{itemize}
		\item ... and not so good ones:
		      \begin{itemize}
			      \item {\texttt{Never use monospace for normal text}}
			      \item {\frakfamily Gothic, calligraphic or weird fonts: should always: be
			            avoided}
		      \end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Look}
	\begin{itemize}
		\item To insert a final slide with the title and final thanks, use \verb|\backmatter|.
		      \begin{itemize}
			      \item The title also appears in footlines along with the author name, you can change this text with \verb|\footlinepayoff|
			      \item You can remove the title from the final slide with \verb|\backmatter[notitle]|
		      \end{itemize}
		\item The aspect ratio defaults to 16:9, and you should not change it to 4:3
		      for old projectors as it is inherently impossible to perfectly convert a
		      16:9 presentation to 4:3 one; spacings \emph{will} break
		      \begin{itemize}
			      \item The \texttt{aspectratio} argument to the \texttt{beamer} class is
			            overridden by the SINTEF theme
			      \item If you \emph{really} know what you are doing, check the package
			            code and look for the \texttt{geometry} class.
		      \end{itemize}
	\end{itemize}
\end{frame}

\section{Summary}

\begin{frame}
	\frametitle{Good Luck!}
	\begin{itemize}
		\item Enough for an introduction! You should know enough by now
		\item If you have corrections or suggestions,
		      \hrefcol{mailto:andrea@gasparini.cloud}{send them to me!}
	\end{itemize}
\end{frame}

\backmatter
\end{document}

### 2023-08-16 docs/presentation/presentation.tex
Update presentation

@@
 \usepackage{amsfonts,amsmath,oldgerm}
+\usepackage{bytefield}
 \usetheme{sintef}
-
-\newcommand{\testcolor}[1]{\colorbox{#1}{\textcolor{#1}{test}}~\texttt{#1}}
-
 \usefonttheme[onlymath]{serif}
-
 \titlebackground*{assets/background}
 
-\newcommand{\hrefcol}[2]{\textcolor{cyan}{\href{#1}{#2}}}
+\newcommand{\rot}[1]{\rotatebox{90}{#1}}
 
@@
 
+\setbeameroption{show notes}
+
 \begin{document}
@@
 
-\begin{frame}
-
-	This template is based on \hrefcol{https://www.overleaf.com/latex/templates/sapienza-presentation/yyksmrskffhk}{Sapienza Presentation} from \hrefcol{mailto:andrea@gasparini.cloud}{Andrea Gasparini}
-
-	\vspace{\baselineskip}
-
-	The above template is a based on \hrefcol{https://www.overleaf.com/latex/templates/sintef-presentation/jhbhdffczpnx}{SINTEF Presentation} from \hrefcol{mailto:federico.zenith@sintef.no}{Federico Zenith} and its derivation \hrefcol{https://github.com/TOB-KNPOB/Beamer-LaTeX-Themes}{Beamer-LaTeX-Themes} from Liu Qilong
-
-	\vspace{\baselineskip}
-
-	In the following you find a brief introduction on how to use \LaTeX\ and the beamer package to prepare slides, based on the one written by \hrefcol{mailto:federico.zenith@sintef.no}{Federico Zenith} for \hrefcol{https://www.overleaf.com/latex/templates/sintef-presentation/jhbhdffczpnx}{SINTEF Presentation}
-
-	% This template is released under \hrefcol{https://creativecommons.org/licenses/by-nc/4.0/legalcode}{Creative Commons CC BY 4.0} license
-
-\end{frame}
+% \begin{frame}
+%
+% 	This template is based on \hrefcol{https://www.overleaf.com/latex/templates/sapienza-presentation/yyksmrskffhk}{Sapienza Presentation} from \hrefcol{mailto:andrea@gasparini.cloud}{Andrea Gasparini}
+%
+% 	\vspace{\baselineskip}
+%
+% 	The above template is a based on \hrefcol{https://www.overleaf.com/latex/templates/sintef-presentation/jhbhdffczpnx}{SINTEF Presentation} from \hrefcol{mailto:federico.zenith@sintef.no}{Federico Zenith} and its derivation \hrefcol{https://github.com/TOB-KNPOB/Beamer-LaTeX-Themes}{Beamer-LaTeX-Themes} from Liu Qilong
+%
+% 	\vspace{\baselineskip}
+%
+% 	In the following you find a brief introduction on how to use \LaTeX\ and the beamer package to prepare slides, based on the one written by \hrefcol{mailto:federico.zenith@sintef.no}{Federico Zenith} for \hrefcol{https://www.overleaf.com/latex/templates/sintef-presentation/jhbhdffczpnx}{SINTEF Presentation}
+%
+% 	% This template is released under \hrefcol{https://creativecommons.org/licenses/by-nc/4.0/legalcode}{Creative Commons CC BY 4.0} license
+%
+% \end{frame}
 
@@
 
-\begin{frame}{Beamer for SINTEF slides}
+\begin{frame}{Web Data Interchange Formats}
 	\begin{itemize}
-		\item We assume you can use \LaTeX; if you cannot,
-		      \hrefcol{http://en.wikibooks.org/wiki/LaTeX/}{you can learn it here}
-		\item Beamer is one of the most popular and powerful document
-		      classes for presentations in \LaTeX
-		\item Beamer has also a detailed
-		      \hrefcol{http://www.ctan.org/tex-archive/macros/latex/contrib/beamer/doc/beameruserguide.pdf}{user
-			      manual}
-		\item Here we will present only the most basic features to get you up to speed
+		\item Human-Readable and Self-Describing
+		      \note[item]{Most formats on the web are human-readable and self-describing}
+		      \note[item]{Self describing means they include metadata to describe themselves, and don't require an external schema}
+		      \note[item]{Human readable means plain text, ASCII or utf-8}
+		\item Most common: JSON and XML
+		      \note[item]{The most common of these formats is JSON and XML}
+		\item Alternatives: CSV, MessagePack, CBOR
+		      \note[item]{There are other human-readable formats such as CSV}
+		      \note[item]{And others, such as CBOR (concise binary object representation), and MessagePack that are still self-describing, but use a binary format instead}
 	\end{itemize}
@@
 
-\begin{frame}{Beamer vs. PowerPoint}
-	Compared to PowerPoint, using \LaTeX\ is better because:
+\begin{frame}{Shortcomings of Current Formats}
 	\begin{itemize}
-		\item It is not What-You-See-Is-What-You-Get, but
-		      What-You-\emph{Mean}-Is-What-You-Get:\\
-		      you write the content, the computer does the typesetting
-		\item Produces a \texttt{pdf}: no problems with fonts, formulas,
-		      program versions
-		\item Easier to keep consistent style, fonts, highlighting, etc.
-		\item Math typesetting in \TeX\ is the best:
-		      \begin{equation*}
-			      \mathrm{i}\,\hslash\frac{\partial}{\partial t} \Psi(\mathbf{r},t) =
-			      -\frac{\hslash^2}{2\,m}\nabla^2\Psi(\mathbf{r},t)
-			      + V(\mathbf{r})\Psi(\mathbf{r},t)
-		      \end{equation*}
+		\item Repeated sending of metadata.
+		\item Inefficiency of plain text.
+		      \note[item]{Expensive conversion between plain text and binary}
+		      \note[item]{Requires more bytes to store the same data}
+		\item Use of big-endian.
+		      \note[item]{Most modern processors (x86 and ARM) are little-endian}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Improvements}
+	\begin{itemize}
+		\item When possible, send metadata only a single time (mainly for homogenous lists).
+		\item Allow sending metadata to be optional, such as if a client already has it.
+		      \note[item]{Keep metadata and data separate, not coupled together}
+		\item Utilize little-endian to reduce conversions.
+		\item Prioritize a simple type system.
+		      \note[item]{Hopefully makes the implementation simpler}
+	\end{itemize}
+\end{frame}
+
+\section{Format Design}
+
+\begin{frame}{Primitives}
+	\begin{itemize}
+		\item Boolean
+		      \note[item]{True of False, one byte}
+		      \note[item]{All integers are little-endian}
+		      \note[item]{Integer types are equivalent to the types in stdint.h C header}
+		\item U8, U16, U32, U64, U128, U256
+		      \note[item]{Unsigned Integer Types}
+		      \begin{itemize}
+			      \item Equivalent to uint*\_t types in C
+		      \end{itemize}
+		\item I8, I16, I32, I64, I128, I256
+		      \note[item]{Signed Integer Types}
+		      \begin{itemize}
+			      \item Equivalent to int*\_t types in C
+		      \end{itemize}
+		\item F32, F64
+		      \note[item]{Floating Point Types}
+		      \begin{itemize}
+			      \item IEEE 754 floating point numbers
+		      \end{itemize}
+		\item Bytes  (Non-UTF-8 Bytes)
+		      \note[item]{String but doesn't have to be UTF-8}
+		\item String (UTF-8 Bytes)
+		      \note[item]{Sends length as u64 followed by UTF-8 bytes}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Primitive Metadata}
+	\center
+	\begin{tabular}{|c|c|}
+		\hline
+		\textbf{Type} & \textbf{Discriminant} \\
+		\hline
+		Boolean       & 0                     \\
+		U8            & 1                     \\
+		U16           & 2                     \\
+		U32           & 3                     \\
+		...           & ...                   \\
+		I256          & 12                    \\
+		F32           & 13                    \\
+		F64           & 14                    \\
+		Bytes         & 15                    \\
+		String        & 16                    \\
+		\hline
+	\end{tabular}
+	\note[item]{Metadata is sent as a single byte discriminant value}
+	\note[item]{The discriminant begin at 0 with Boolean, and increase by 1 for each type}
+	\note[item]{Progress in the same order as the previous slide, with String being the final value}
+\end{frame}
+
+\begin{frame}[fragile]{Primitive Example}
+	\begin{itemize}
+		\item 16-Bit Unsigned Integer: 1024 \\
+		      \begin{bytefield}{24}
+			      \bitbox{8}{Disc} & \bitbox{16}{1024} \\
+			      \bitbox{8}{2} & \bitboxes{8}{04} \\
+		      \end{bytefield}
+		\item String: "hello" \\
+		      \begin{bytefield}[bitheight=\widthof{"Disc"}]{28}
+			      \bitbox{2}{\rot{Disc}} & \bitbox{16}{length} & \bitboxes{2}{hello} \\
+			      \bitboxes{2}{{16} {2} {0} {0} {0} {0} {0} {0} {0} {\rot{104}} {\rot{101}} {\rot{108}} {\rot{108}} {\rot{111}}}
+		      \end{bytefield}
 
@@
 
-\begin{frame}[fragile]{Getting Started}
-	\framesubtitle{Selecting the SINTEF Theme}
-	To start working with \texttt{sintefbeamer}, start a \LaTeX\ document with the
-	preamble:
-	\begin{block}{Minimum SINTEF Beamer Document}
-		\verb|\documentclass{beamer}|\\
-		\verb|\usetheme{sintef}|\\
-		\verb|\begin{document}|\\
-		\verb|\begin{frame}{Hello, world!}|\\
-		\verb|\end{frame}|\\
-		\verb|\end{document}|\\
-	\end{block}
-\end{frame}
-
-\begin{frame}[fragile]{Title page}
-	To set a typical title page, you call some commands in the preamble:
-	\begin{block}{The Commands for the Title Page}
-		\begin{verbatim}
-\title{Sample Title}
-\subtitle{Sample subtitle}
-\author{First Author, Second Author}
-\date{\today} % Can also be (ab)used for conference name &c.
-\end{verbatim}
-	\end{block}
-	You can then write out the title page with \verb|\maketitle|.
-
-	To set a \textbf{background image} use the \verb|\titlebackground| command
-	before \verb|\maketitle|; its only argument is the name (or path) of a graphic
-	file.
-
-	If you use the \textbf{starred version} \verb|\titlebackground*|, the image
-	will be clipped to a split view on the right side of the title slide.
-
-\end{frame}
-
-\begin{frame}[fragile]{Writing a Simple Slide}
-	\framesubtitle{It's really easy!}
-	\begin{itemize}[<+->]
-		\item A typical slide has bulleted lists
-		\item These can be uncovered in sequence
-	\end{itemize}
-	\begin{block}{Code for a Page with an Itemised List}<+->
-		\begin{verbatim}
-\begin{frame}{Writing a Simple Slide}
-\framesubtitle{It's really easy!}
-\begin{itemize}[<+->]
-\item A typical slide has bulleted lists
-\item These can be uncovered in sequence
-\end{itemize}\end{frame}
-\end{verbatim}
-	\end{block}
-\end{frame}
-
-\section{Personalization}
-
-\footlinecolor{stevensmediumgold}
-\begin{frame}[fragile]{Changing Slide Style}
+\begin{frame}{Vectors (Homogenous Lists)}
 	\begin{itemize}
-		\item You can select the white or \textit{maincolor} \textbf{slide style} \emph{in the
-			      preamble} with \verb|\themecolor{white}| (default) or \verb|\themecolor{main}|
-		      \begin{itemize}
-			      \item You should \emph{not} change these within the document: Beamer does
-			            not like it
-			      \item If you \emph{really} must, you may have to add
-			            \verb|\usebeamercolor[fg]{normal text}| in the slide
-		      \end{itemize}
-		\item You can change the \textbf{footline colour} with
-		      \verb|\footlinecolor{color}|
-		      \begin{itemize}
-			      \item Place the command \emph{before} a new \verb|frame|
-			      \item There are four ``official'' colors:
-			            \testcolor{maincolor}, \testcolor{sintefyellow},
-			            \testcolor{sintefgreen}, \testcolor{sintefdarkgreen}
-			      \item Default is no footline; you can restore it with
-			            \verb|\footlinecolor{}|
-			      \item Others may work, but no guarantees!
-			      \item Should \emph{not} be used with the \verb|maincolor| theme!
-		      \end{itemize}
+		\item Homogenous list of values that has a known, variable length.
+		\item Length is sent as a part of the type, not in metadata.
+		      \note[item]{Length is an u64, the same as String and Bytes}
+		\item Length is sent, followed by the elements.
+		      \note[item]{No type info about elements, that's in metadata}
+		\item Inner type Information is \emph{not} sent.
 	\end{itemize}
@@
 
-\begin{frame}[fragile]{Blocks}
-	\begin{columns}
-		\begin{column}{0.3\textwidth}
-			\begin{block}{Standard Blocks}
-				These have a color coordinated with the footline (and grey in the blue theme)
-				\begin{verbatim}
-\begin{block}{title}
-content...
-\end{block}
-\end{verbatim}
-			\end{block}
-		\end{column}
-		\begin{column}{0.7\textwidth}
-			\begin{colorblock}[black]{stevenslightorange}{Colour Blocks}
-				Similar to the ones on the left, but you pick the colour. Text will be white by
-				default, but you may set it with an optional argument.
-				\small
-				\begin{verbatim}
-\begin{colorblock}[black]{sinteflightgreen}{title}
-content...
-\end{colorblock}
-\end{verbatim}
-			\end{colorblock}
-			The ``official'' colours of colour blocks are: \testcolor{sinteflilla},
-			\testcolor{maincolor}, \testcolor{sintefdarkgreen}, and
-			\testcolor{sintefyellow}.
-		\end{column}
-	\end{columns}
-\end{frame}
-
-\footlinecolor{}
-\begin{frame}[fragile]{Using Colours}
-	\begin{itemize}[<alert@2>]
-		\item You can use colours with the
-		      \verb|\textcolor{<color name>}{text}| command
-		\item The colours are defined in the \texttt{sintefcolor} package:
+\begin{frame}{Vector Metadata}
 	\begin{itemize}
-			      \item Primary colours: \testcolor{maincolor} and its sidekick
-			            \testcolor{stevensgray}
-			      \item Secondary Colors are: \testcolor{stevensdarkgray}, \testcolor{stevensdarkblue},\testcolor{stevensmediumblue}, \testcolor{stevensmediumorange}, \testcolor{stevensmediumgold}
-			      \item Light colors are:
-			            \testcolor{stevenslightgray},\testcolor{stevenslightblue}, \testcolor{stevenslightorange}, \testcolor{stevenslightgold}
-			      \item Other colors are \testcolor{sinteflightgreen},
-			            \testcolor{sintefgreen}, \testcolor{sintefdarkgreen}
-			      \item Additional colours: \testcolor{sintefyellow}, \testcolor{sintefred},
-			            \testcolor{sinteflilla}
-			            \begin{itemize}
-				            \item These may be shaded---see the \verb|sintefcolor| documentation or
-				                  the \hrefcol{https://sintef.sharepoint.com/sites/stottetjenester/%
-					                  kommunikasjon/grafisk-profil-new/Sider/default.aspx}{SINTEF profile
-					                  manual}
-			            \end{itemize}
-		      \end{itemize}
-		\item Do \emph{not} abuse colours: \verb|\emph{}| is usually enough
-		\item Use \verb|\alert{}| to bring the \alert<2->{focus} somewhere
-		\item<2- | alert@2> If you highlight too much, you don't highlight at all!
+		\item Single byte discriminant value is sent (String + 1).
+		\item Followed by internal type metadata (can continue recursively).
+		\item Length is \emph{not} sent.
 	\end{itemize}
@@
 
-\begin{frame}[fragile]{Adding images}
-	\begin{columns}
-		\begin{column}{0.7\textwidth}
-			Adding images works like in normal \LaTeX:
-			\begin{block}{Code for Adding Images}
-				\begin{verbatim}
-\usepackage{graphicx}
-% ...
-\includegraphics[width=\textwidth]
-{assets/logo_RGB}
-\end{verbatim}
-			\end{block}
-		\end{column}
-		\begin{column}{0.3\textwidth}
-			\includegraphics[width=\textwidth]
-			{assets/logo_RGB}
-		\end{column}
-	\end{columns}
+\begin{frame}[fragile]{Vector Example}
+	\begin{center}
+		\textbf{Vector of 16-Bit Unsigned Integers:} [42, 1024]
+
+		DV = Discriminant of a Vector
+
+		D16 = Discriminant of a 16-Bit Unsigned Integer
+
+		\begin{bytefield}{26}
+			\bitbox{2}{\rot{DV}} & \bitbox{2}{\rot{D16}} & \bitbox{16}{length} & \bitboxes{4}{{42} {1024}} \\
+			\bitboxes{2}{{17} {2} {2} {0} {0} {0} {0} {0} {0} {0} {42} {0} {0} {4}}
+		\end{bytefield}
+	\end{center}
 \end{frame}
 
-\begin{frame}[fragile]{Splitting in Columns}
-	Splitting the page is easy and common;
-	typically, one side has a picture and the other text:
-	\begin{columns}
-		\begin{column}{0.6\textwidth}
-			This is the first column
-		\end{column}
-		\begin{column}{0.3\textwidth}
-			And this the second
-		\end{column}
-	\end{columns}
-	\begin{block}{Column Code}
-		\begin{verbatim}
-\begin{columns}
-\begin{column}{0.6\textwidth}
-  This is the first column
-\end{column}
-\begin{column}{0.3\textwidth}
-  And this the second
-\end{column}
-% There could be more!
-\end{columns}
-\end{verbatim}
-	\end{block}
-\end{frame}
-
-\begin{chapter}[assets/background_negative]{}{Special Slides}
+\begin{frame}{Structs (Aggregate Types)}
 	\begin{itemize}
-	 \item Chapter slides
-	 \item Side-picture slides
- \end{itemize}
-\end{chapter}
-
-\footlinecolor{sintefred}
-\begin{frame}[fragile]{Chapter slides}
-	\begin{itemize}
-		\item Similar to \verb|frame|s, but with a few more options
-		\item Opened with \verb|\begin{chapter}[<image>]{<color>}{<title>}|
-		\item Image is optional, colour and title are mandatory
-		\item There are seven ``official'' colours: \testcolor{maincolor},
-		      \testcolor{sintefdarkgreen}, \testcolor{sintefgreen},
-		      \testcolor{sinteflightgreen}, \testcolor{sintefred}, \testcolor{sintefyellow},
-		      \testcolor{sinteflilla}.
-		      \begin{itemize}
-			      \item Strangely enough, these are \emph{more} than the official colours
-			            for the footline.
-			      \item It may still be a nice touch to change the footline of following
-			            slides to the same color of a chapter slide. Your choice.
-		      \end{itemize}
-		\item Otherwise, \verb|chapter| behaves just like \verb|frame|.
+		\item item
 	\end{itemize}
@@
 
-\begin{sidepic}{assets/background_alternative}{Side-Picture Slides}
+\begin{frame}{Struct Metadata}
 	\begin{itemize}
-		\item Opened with \texttt{$\backslash$begin\{sidepic\}\{<image>\}\{<title>\}}
-		\item Otherwise, \texttt{sidepic} works just like \texttt{frame}
-	\end{itemize}
-\end{sidepic}
-
-\footlinecolor{}
-\begin{frame}
-	\frametitle{Fonts}
-	\begin{itemize}
-		\item The paramount task of fonts is being readable
-		\item There are good ones...
-		      \begin{itemize}
-			      \item {\textrm{Use serif fonts only with high-definition projectors}}
-			      \item {\textsf{Use sans-serif fonts otherwise (or if you simply prefer
-				            them)}}
-		      \end{itemize}
-		\item ... and not so good ones:
-		      \begin{itemize}
-			      \item {\texttt{Never use monospace for normal text}}
-			      \item {\frakfamily Gothic, calligraphic or weird fonts: should always: be
-			            avoided}
-		      \end{itemize}
+		\item item
 	\end{itemize}
@@
 
-\begin{frame}[fragile]{Look}
-	\begin{itemize}
-		\item To insert a final slide with the title and final thanks, use \verb|\backmatter|.
-		      \begin{itemize}
-			      \item The title also appears in footlines along with the author name, you can change this text with \verb|\footlinepayoff|
-			      \item You can remove the title from the final slide with \verb|\backmatter[notitle]|
-		      \end{itemize}
-		\item The aspect ratio defaults to 16:9, and you should not change it to 4:3
-		      for old projectors as it is inherently impossible to perfectly convert a
-		      16:9 presentation to 4:3 one; spacings \emph{will} break
-		      \begin{itemize}
-			      \item The \texttt{aspectratio} argument to the \texttt{beamer} class is
-			            overridden by the SINTEF theme
-			      \item If you \emph{really} know what you are doing, check the package
-			            code and look for the \texttt{geometry} class.
-		      \end{itemize}
-	\end{itemize}
-\end{frame}
+\section{Results}
 
-\section{Summary}
-
-\begin{frame}
-	\frametitle{Good Luck!}
-	\begin{itemize}
-		\item Enough for an introduction! You should know enough by now
-		\item If you have corrections or suggestions,
-		      \hrefcol{mailto:andrea@gasparini.cloud}{send them to me!}
-	\end{itemize}
-\end{frame}
-
-\backmatter
+\backmatter[notitle]
 \end{document}

### 2023-08-16 docs/presentation/presentation.tex
Update presentation and paper

@@
 \usepackage{bytefield}
+\usepackage{listings}
+
+\lstdefinestyle{rust}{
+	tabsize=2
+}
+\lstset{style=rust}
+
 \usetheme{sintef}
@@
 
-\setbeameroption{show notes}
+% \setbeameroption{show notes}
 
@@
 \maketitle
-
-% \begin{frame}
-%
-% 	This template is based on \hrefcol{https://www.overleaf.com/latex/templates/sapienza-presentation/yyksmrskffhk}{Sapienza Presentation} from \hrefcol{mailto:andrea@gasparini.cloud}{Andrea Gasparini}
-%
-% 	\vspace{\baselineskip}
-%
-% 	The above template is a based on \hrefcol{https://www.overleaf.com/latex/templates/sintef-presentation/jhbhdffczpnx}{SINTEF Presentation} from \hrefcol{mailto:federico.zenith@sintef.no}{Federico Zenith} and its derivation \hrefcol{https://github.com/TOB-KNPOB/Beamer-LaTeX-Themes}{Beamer-LaTeX-Themes} from Liu Qilong
-%
-% 	\vspace{\baselineskip}
-%
-% 	In the following you find a brief introduction on how to use \LaTeX\ and the beamer package to prepare slides, based on the one written by \hrefcol{mailto:federico.zenith@sintef.no}{Federico Zenith} for \hrefcol{https://www.overleaf.com/latex/templates/sintef-presentation/jhbhdffczpnx}{SINTEF Presentation}
-%
-% 	% This template is released under \hrefcol{https://creativecommons.org/licenses/by-nc/4.0/legalcode}{Creative Commons CC BY 4.0} license
-%
-% \end{frame}
 
@@
 
-\begin{frame}{Improvements}
+\begin{frame}{Improvements For a New Format}
 	\begin{itemize}
@@
 			      \item IEEE 754 floating point numbers
+			      \item Equivalent to float and double in C
 		      \end{itemize}
@@
 		      \note[item]{No type info about elements, that's in metadata}
-		\item Inner type Information is \emph{not} sent.
+		\item Inner type Information is \emph{not} sent as a part of the value.
 	\end{itemize}
@@
 		\item Followed by internal type metadata (can continue recursively).
-		\item Length is \emph{not} sent.
+		\item Length is \emph{not} sent as a part of the metadata.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item item
+		\item	An aggregate type containing a name as well as named fields.
+		\item May not contain duplicate field names.
+		\item Fields are sent in sequence in the order they are listed in metadata.
+		\item Serialized value does \emph{not} include:
+		      \begin{itemize}
+			      \item Struct name.
+			      \item Number of fields.
+			      \item Field names.
+			      \item Field types.
+		      \end{itemize}
 	\end{itemize}
@@
 	\begin{itemize}
-		\item item
+		\item Single byte discriminant value is sent (Vector + 1).
+		\item Followed by Struct name.
+		\item Next, number of fields as an unsigned 16-bit integer.
+		      \note[item]{A 16-bit integer is used because it's very unlikely for a struct of more than 65 thousand fields to be necessary.}
+		\item Last, pairs of field names and types.
+		      \begin{itemize}
+			      \item May continue recursively with nested types of Structs or Vectors.
+		      \end{itemize}
+		\item All names are sent with the same format as primitive Strings.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}[fragile]{Struct Metadata Example}
+	\begin{columns}
+		\begin{column}{0.65\textwidth}
+			\note[item]{ DS = Discriminant of a Struct
+
+				DX = Discriminant of field "first"
+
+				DY = Discriminant of field "last"}
+			\center
+			\begin{bytefield}{32}
+				\bitbox{2}{\rot{DS}} & \bitbox{16}{struct name length} & \bitboxes{2}{Point} & \bitbox{4}{fields} \\
+				\bitboxes{2}{{18} {5} {0} {0} {0} {0} {0} {0} {0} {80} {\rot{111}} {\rot{105}} {\rot{110}} {\rot{116}} {2} {0}} \\
+			\end{bytefield}
+			\begin{bytefield}{32}
+				\bitbox{16}{first name length} & \bitboxes{2}{first} & \bitbox{2}{\rot{DX}} \\
+				\bitboxes{2}{{5} {0} {0} {0} {0} {0} {0} {0} {\rot{102}} {\rot{105}} {\rot{114}} {\rot{115}} {\rot{116}} {7}} \\
+			\end{bytefield}
+			\begin{bytefield}{32}
+				\bitbox{16}{last name length} & \bitboxes{2}{last} & \bitbox{2}{\rot{DY}} \\
+				\bitboxes{2}{{4} {0} {0} {0} {0} {0} {0} {0} {\rot{108}} {\rot{97}} {\rot{115}} {\rot{116}} {7}}
+			\end{bytefield}
+		\end{column}
+		\begin{column}{0.35\textwidth}
+			\begin{lstlisting}
+struct Point {
+	first: i8,
+	last: i8
+}
+\end{lstlisting}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}[fragile]{Struct Value Example}
+	\begin{columns}
+		\begin{column}{0.5\textwidth}
+			\center
+			\begin{bytefield}{12}
+				\bitbox{6}{first} & \bitbox{6}{last} \\
+				\bitbox{6}{42} & \bitbox{6}{251} \\
+			\end{bytefield}
+		\end{column}
+		\begin{column}{0.5\textwidth}
+			\center
+			\begin{lstlisting}
+let my_point = Point {
+	first: 42,
+	last: -5,
+};
+			\end{lstlisting}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\section{Evaluation}
+
+\begin{frame}{Tested Formats}
+	\begin{itemize}
+		\item CSV
+		\item MessagePack
+		\item CBOR
+		\item JSON
+		\item XML
 	\end{itemize}
@@
 
+\begin{frame}{Average Time and Bytes Read}
+	\center
+	\begin{tabular}{|c|c|c|}
+		\hline
+		\textbf{Format} & \textbf{Avg Time (ms)} & \textbf{Bytes Read} \\
+		\hline
+		CSV             & 18.93                  & 16,411              \\
+		MessagePack     & 11.22                  & 15,565              \\
+		CBOR            & 16.95                  & 25,507              \\
+		JSON            & 21.91                  & 31,180              \\
+		XML             & 21.87                  & 43,699              \\
+		XBF             & 11.32                  & 14,686              \\
+		\hline
+	\end{tabular}
+\end{frame}
+
+\begin{frame}{Overhead}
+	\center
+	\begin{tabular}{|c|c|c|}
+		\hline
+		\textbf{Format} & \textbf{Overhead (bytes)} & \textbf{Percent Overhead} \\
+		\hline
+		CSV             & 1,823                     & 11.1                      \\
+		MessagePack     & 977                       & 6.22                      \\
+		CBOR            & 10,919                    & 42.8                      \\
+		JSON            & 16,592                    & 53.2                      \\
+		XML             & 29,111                    & 66.6                      \\
+		XBF             & 98                        & 0.67                      \\
+		\hline
+	\end{tabular}
+\end{frame}
+
 \backmatter[notitle]

### 2023-08-18 docs/presentation/presentation.tex
Update paper and presentation

@@
 
-% \setbeameroption{show notes}
+\setbeameroption{show notes}
 
@@
 		      \note[item]{Most modern processors (x86 and ARM) are little-endian}
+		      \note[item]{Both CBOR and MessagePack are big-endian}
 	\end{itemize}
+	% XML from paper here
 \end{frame}
+
+% possibly add three slides explaining each of these
+% 1: XML being huge
+% 2: plain text example from paper
+% 3: 1027 as u32 0 4 3 0 -> 0 3 4 0
 
@@
 		      \note[item]{Keep metadata and data separate, not coupled together}
-		\item Utilize little-endian to reduce conversions.
+		\item Utilize little-endian to eliminate endian conversions on common architectures.
 		\item Prioritize a simple type system.
@@
 		      \note[item]{All integers are little-endian}
-		      \note[item]{Integer types are equivalent to the types in stdint.h C header}
+		      \note[item]{Integer types are equivalent to the types in stdint.h C headers}
 		\item U8, U16, U32, U64, U128, U256
@@
 	\end{tabular}
+	% two columsn for each half of the list
 	\note[item]{Metadata is sent as a single byte discriminant value}
@@
 	\end{columns}
+	% add explanation for DS, DX, DY
 \end{frame}

### 2023-08-18 docs/presentation/presentation.tex
Add visual examples for inefficiencies
Update presentation

@@
 \title{Implementing XBF: An Efficient Self-Describing Binary Format}
-% \subtitle{Using \LaTeX\ to prepare slides}
-% \course{Master's Degree in Computer Science}
 \author{\href{mailto:dkrautha@pm.me}{David Krauthamer}}
-% \IDnumber{1234567}
 
@@
 
-% possibly add three slides explaining each of these
-% 1: XML being huge
-% 2: plain text example from paper
-% 3: 1027 as u32 0 4 3 0 -> 0 3 4 0
+\begin{frame}[fragile]{Repeated Metadata}
+	\begin{columns}
+		\begin{column}{0.5\textwidth}
+			\note[item]{XML example where a huge percent of the actual data sent is metadata}
+			\begin{lstlisting}[language=XML]
+<person>
+	<name>John Jackson</name>
+	<age>25</age>
+</person>
+\end{lstlisting}
+		\end{column}
+		\begin{column}{0.5\textwidth}
+			\begin{table}
+				\caption{XML Space Utilization}
+				\center
+				\begin{tabular}{|c|c|c|}
+					\hline
+					\textbf{} & \textbf{Bytes Used} & \textbf{Percent} \\
+					\hline
+					Metadata  & 45                  & 76\%             \\
+					Data      & 14                  & 24\%             \\
+					\hline
+					Total     & 59                  & 100              \\
+					\hline
+				\end{tabular}
+			\end{table}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}[fragile]{Plain Text Example}
+	\begin{table}
+		\caption{Sending A 32-bit Unsigned Integer}
+		\begin{center}
+			\begin{tabular}{|c|c|c|}
+				\hline
+				\textbf{Number} & \textbf{ASCII Bytes} & \textbf{Binary Bytes} \\
+				\hline
+				1               & 1                    & 4                     \\
+				9999            & 4                    & 4                     \\
+				7654321         & 7                    & 4                     \\
+				\hline
+			\end{tabular}
+		\end{center}
+	\end{table}
+\end{frame}
+
+\begin{frame}[fragile]{Big Endian Example: 32-bit Unsigned Integer}
+	\begin{center}
+		\textbf{Number To Send: 1027} \\
+	\end{center}
+	\begin{columns}
+		\begin{column}{0.3\textwidth}
+			\begin{bytefield}{16}
+				\wordbox{1}{Sender (Little Endian)} \\
+				\bitboxes{4}{{3} {4} {0} {0}} \\
+			\end{bytefield}
+		\end{column}
+		\begin{column}{0.3\textwidth}
+			\begin{bytefield}{16}
+				\wordbox{1}{On the Wire (Big Endian)} \\
+				\bitboxes{4}{{0} {0} {4} {3}} \\
+			\end{bytefield}
+		\end{column}
+		\begin{column}{0.3\textwidth}
+			\begin{bytefield}{16}
+				\wordbox{1}{Recipient (Little Endian)} \\
+				\bitboxes{4}{{3} {4} {0} {0}} \\
+			\end{bytefield}
+		\end{column}
+	\end{columns}
+\end{frame}
 
@@
 \begin{frame}{Primitive Metadata}
+	\begin{columns}
+		\begin{column}{0.5\textwidth}
 			\center
@@
 				U32           & 3                     \\
-		...           & ...                   \\
+				U64           & 4                     \\
+				U128          & 5                     \\
+				U256          & 6                     \\
+				\hline
+			\end{tabular}
+		\end{column}
+		\begin{column}{0.5\textwidth}
+			\center
+			\begin{tabular}{|c|c|}
+				\hline
+				\textbf{Type} & \textbf{Discriminant} \\
+				\hline
+				I8            & 7                     \\
+				I16           & 8                     \\
+				I32           & 9                     \\
+				I64           & 10                    \\
+				I128          & 11                    \\
 				I256          & 12                    \\
@@
 			\end{tabular}
-	% two columsn for each half of the list
+		\end{column}
+	\end{columns}
 	\note[item]{Metadata is sent as a single byte discriminant value}
 	\note[item]{The discriminant begin at 0 with Boolean, and increase by 1 for each type}
-	\note[item]{Progress in the same order as the previous slide, with String being the final value}
 \end{frame}
@@
 		\begin{column}{0.65\textwidth}
-			\note[item]{ DS = Discriminant of a Struct
-
-				DX = Discriminant of field "first"
-
-				DY = Discriminant of field "last"}
 			\center
@@
 			\begin{bytefield}{32}
-				\bitbox{16}{first name length} & \bitboxes{2}{first} & \bitbox{2}{\rot{DX}} \\
+				\bitbox{16}{first name length} & \bitboxes{2}{first} & \bitbox{2}{\rot{DF}} \\
 				\bitboxes{2}{{5} {0} {0} {0} {0} {0} {0} {0} {\rot{102}} {\rot{105}} {\rot{114}} {\rot{115}} {\rot{116}} {7}} \\
@@
 			\begin{bytefield}{32}
-				\bitbox{16}{last name length} & \bitboxes{2}{last} & \bitbox{2}{\rot{DY}} \\
+				\bitbox{16}{last name length} & \bitboxes{2}{last} & \bitbox{2}{\rot{DL}} \\
 				\bitboxes{2}{{4} {0} {0} {0} {0} {0} {0} {0} {\rot{108}} {\rot{97}} {\rot{115}} {\rot{116}} {7}}
@@
 \end{lstlisting}
+			\hfill \break
+			DS = Struct Discriminant
+
+			DF = Field "first" discriminant
+
+			DL = Field "last" discriminant
 		\end{column}
 	\end{columns}
-	% add explanation for DS, DX, DY
 \end{frame}

### 2023-08-19 docs/presentation/presentation.tex
Update presentation

@@
 
+% pympress notes page config
+\iffalse
+	\setbeamertemplate{note page}
+	{
+		\insertslideintonotes{0.01}
+		\rule{\textwidth}{0.1pt}
+		%\color{blue} \scriptsize
+		\insertnote
+	}
+	\setbeameroption{show notes on second screen}
+\fi
+
+% adjust gigantic tabs in code blocks, not actually rust format
 \lstdefinestyle{rust}{
@@
 
+% needed this a bunch so made a macro for myself
 \newcommand{\rot}[1]{\rotatebox{90}{#1}}
@@
 \author{\href{mailto:dkrautha@pm.me}{David Krauthamer}}
-
-\setbeameroption{show notes}
 
@@
 
-\begin{frame}{Primitives}
+\begin{frame}{Primitives and Primitive Metadata}
+	\begin{columns}
+		\begin{column}{0.5\textwidth}
 			\begin{itemize}
@@
 			\end{itemize}
-\end{frame}
-
-\begin{frame}{Primitive Metadata}
-	\begin{columns}
-		\begin{column}{0.5\textwidth}
+		\end{column}
+		\begin{column}{0.25\textwidth}
 			\center
@@
 				\hline
-				\textbf{Type} & \textbf{Discriminant} \\
+				\textbf{Type} & \textbf{Value} \\
 				\hline
@@
 		\end{column}
-		\begin{column}{0.5\textwidth}
+		\begin{column}{0.25\textwidth}
 			\center
@@
 	\end{columns}
-	\note[item]{Metadata is sent as a single byte discriminant value}
-	\note[item]{The discriminant begin at 0 with Boolean, and increase by 1 for each type}
 \end{frame}
@@
 \begin{frame}{Vectors (Homogenous Lists)}
+	Values
 	\begin{itemize}
-		\item Homogenous list of values that has a known, variable length.
+		\item Homogenous list of values that has a known length.
 		\item Length is sent as a part of the type, not in metadata.
-		      \note[item]{Length is an u64, the same as String and Bytes}
-		\item Length is sent, followed by the elements.
-		      \note[item]{No type info about elements, that's in metadata}
 		\item Inner type Information is \emph{not} sent as a part of the value.
 	\end{itemize}
-\end{frame}
 
-\begin{frame}{Vector Metadata}
+	Metadata
 	\begin{itemize}
@@
 	\begin{center}
-		\textbf{Vector of 16-Bit Unsigned Integers:} [42, 1024]
+		Vector of 16-Bit Unsigned Integers: [42, 1024] \vfill \break
 
@@
 
-		D16 = Discriminant of a 16-Bit Unsigned Integer
+		D16 = Discriminant of a 16-Bit Unsigned Integer \vfill \break
 
@@
 
-\begin{frame}{Tested Formats}
+\begin{frame}{Encoder and Decoder Reference Implementation}
 	\begin{itemize}
-		\item CSV
-		\item MessagePack
-		\item CBOR
-		\item JSON
-		\item XML
+		\item Written in Rust.
+		      \note[item]{systems level programming language similar to C and C++, strong memory safety guarantees}
+		\item Extensively Tested.
+		      \note[item]{100 percent code coverage}
+		      \note[item]{not going to talk a ton about the reference implementation}
+		\item Majority of time was spent on this implementation.
 	\end{itemize}
+\end{frame}
+
+\begin{frame}{Performance Test}
+	\begin{columns}
+		\begin{column}{0.5\textwidth}
+			Server
+			\begin{itemize}
+				\item Downloads and parses 1 year of Sony stock history.
+				\item 
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.5\textwidth}
+				\item For each data format the client:
+				      \begin{itemize}
+					      \item Client sends 100 requests to the server.
+					      \item Records average time to receive a response.
+					      \item Records number of bytes read.
+				      \end{itemize}
+		\end{column}
+	\end{columns}
 \end{frame}
@@
 	\center
-	\begin{tabular}{|c|c|c|}
+	\begin{tabular}{|c|c|c|c|c|}
 		\hline
-		\textbf{Format} & \textbf{Avg Time (ms)} & \textbf{Bytes Read} \\
+		\textbf{Format} & \textbf{Avg Time (ms)} & \textbf{Bytes Read} & \textbf{Overhead (bytes)} & \textbf{Percent Overhead} \\ 
 		\hline                                                                                                                    
-		CSV             & 18.93                  & 16,411              \\
-		MessagePack     & 11.22                  & 15,565              \\
-		CBOR            & 16.95                  & 25,507              \\
-		JSON            & 21.91                  & 31,180              \\
-		XML             & 21.87                  & 43,699              \\
-		XBF             & 11.32                  & 14,686              \\
+		CSV             & 18.93                  & 16,411              & 1,823                     & 11.1                      \\ 
+		MessagePack     & 11.22                  & 15,565              & 977                       & 6.22                      \\ 
+		CBOR            & 16.95                  & 25,507              & 10,919                    & 42.8                      \\ 
+		JSON            & 21.91                  & 31,180              & 16,592                    & 53.2                      \\ 
+		XML             & 21.87                  & 43,699              & 29,111                    & 66.6                      \\ 
+		XBF             & 11.32                  & 14,686              & 98                        & 0.67                      \\ 
 		\hline

### 2023-08-20 docs/paper/paper.tex
Update presentation and paper

@@
 In this particular example, 14 characters of the 59 total characters actually contain the desired data, the rest are just describing the metadata.
-
-% An additional optimization that could be made (but isn't with either of these formats) is deduplication of the metadata.
-Both JSON and XML include the metadata for every element sent, even if what is being sent is a homogenous list of elements that all share the same structure.
-%
 
@@
 
-Strings should always be the final value in the list. The value given to strings
-is used by Vectors and Structs to determine what their discriminant value should
+Strings will always be the final value in the list. The value given to strings
+is used by Vectors and Structs to determine what their discriminant value will
 be.
@@
 Vectors are a homogenous list of values that has a known, variable length.
-Vectors should first include their length as an unsigned 64-bit integer (the same as Strings and Bytes), followed by the corresponding number of elements.
+Vectors will first include their length as an unsigned 64-bit integer (the same as Strings and Bytes), followed by the corresponding number of elements.
 The type contained within a Vector is not sent to the client.
@@
 Similarly to primitives, a single byte discriminant value signifying a Vector is incoming is sent.
-This discriminant value should be 1 greater than that of the discriminant value for Strings.
+This discriminant value will be 1 greater than that of the discriminant value for Strings.
 Following this, metadata information for the internal type contained within the Vector will be sent.
@@
 A struct may not contain duplicate field names.
-Should a Struct be sent like this anyway, it should be considered malformed and not be constructed on the receiving end.
+Should a Struct be sent like this anyway, it is considered malformed and may not be constructed on the receiving end.
 Fields of a Struct are sent in sequence in the order they are listed in the Struct's metadata.
-When a struct is serialized it should not send any sort of name information (such as its name or field names), how many fields it has, nor should it send any type information about its fields.
+When a struct is serialized it does not send any sort of name information (such as its name or field names), how many fields it has, nor does it send any type information about its fields.
 That information is carried in the metadata.
@@
 
-A discriminant value should first be sent, similarly to primitives (following the same size requirement). This discriminant value should be 1 greater than that of the discriminant value for Vectors.
+A discriminant value will first be sent, similarly to primitives (following the same size requirement). This discriminant value is 1 greater than that of the discriminant value for Vectors.
 
-Following this, the name of the Struct should be sent, using the same format as primitive strings are sent (unsigned 64-bit length and then the bytes).
+Following this, the name of the Struct will be sent, using the same format as primitive strings are sent (unsigned 64-bit length and then the bytes).
 Next, send the number of fields contained within the Struct as an unsigned 16-bit integer.
 A 16-bit integer is used because it's very unlikely for a struct of more than 65 thousand fields to be necessary.
-Finally, the fields of the Struct should be sent, first the name of the field as a String, then immediately after the metadata for the type of the field.
+Finally, the fields of the Struct will be sent, first the name of the field as a String, then immediately after the metadata for the type of the field.
 This process may continue recursively with nested types of Structs or Vectors.
@@
 The goal is to serialize the variable \texttt{my\_point}, which is an instance of Point with the values 42 and -5 for the members first and last respectively.
-The struct's metadata should be sent as follows:
+The struct's metadata will be sent as follows:
 
@@
 
-The values inside the struct should then be sent in the same order as they are listed in the struct's metadata:
+The values inside the struct will be sent in the same order as they are listed in the struct's metadata:
 
@@
 			\hline
-			\textbf{XBF Type} & \textbf{Rust Type}          & \textbf{C++ Type}                         \\
+			\textbf{XBF Type} & \textbf{Rust Type}          & \textbf{C++ Equivalent Type}              \\
 			\hline
@@
 It makes the same UTF-8 guarantee as a string slice, but has the same properties as a Vec where it keeps track of its capacity, and will reallocate if its capacity is exceeded, similar to a dynamic array.
-As with the decision to use a boxed slice in XbfStruct, a boxed string slice is utilized to save on memory usage, as the metadata for a struct shouldn't need to change, thereby making a capacity field irrelevant.
+As with the decision to use a boxed slice in XbfStruct, a boxed string slice is utilized to save on memory usage, as the metadata for a struct won't need to change, thereby making a capacity field irrelevant.
 
@@
 
+See Table \ref{time_and_bytes} and \ref{overhead} for additional results. Overhead is calculated as native data size recorded by the server subtracted from bytes sent for a format.
+
 \section{Conclusions}

### 2023-08-20 docs/presentation/presentation.tex

@@
 
-\title{Implementing XBF: An Efficient Self-Describing Binary Format}
+\title{Xtensible Binary Format (XBF): An Efficient Self-Describing Binary Format}
 \author{\href{mailto:dkrautha@pm.me}{David Krauthamer}}
@@
 	\begin{itemize}
+		\item Most common: JSON and XML
 		\item Human-Readable and Self-Describing
-		      \note[item]{Most formats on the web are human-readable and self-describing}
-		      \note[item]{Self describing means they include metadata to describe themselves, and don't require an external schema}
-		      \note[item]{Human readable means plain text, ASCII or utf-8}
-		\item Most common: JSON and XML
-		      \note[item]{The most common of these formats is JSON and XML}
 		\item Alternatives: CSV, MessagePack, CBOR
-		      \note[item]{There are other human-readable formats such as CSV}
-		      \note[item]{And others, such as CBOR (concise binary object representation), and MessagePack that are still self-describing, but use a binary format instead}
 	\end{itemize}
@@
 	\begin{itemize}
+		\item Inefficiency of plain text.
 		\item Repeated sending of metadata.
-		\item Inefficiency of plain text.
-		      \note[item]{Expensive conversion between plain text and binary}
-		      \note[item]{Requires more bytes to store the same data}
 		\item Use of big-endian.
-		      \note[item]{Most modern processors (x86 and ARM) are little-endian}
-		      \note[item]{Both CBOR and MessagePack are big-endian}
 	\end{itemize}
-	% XML from paper here
+\end{frame}
+
+\begin{frame}[fragile]{Plain Text Example}
+	\begin{table}
+		\caption{Sending A 32-bit Unsigned Integer}
+		\begin{center}
+			\begin{tabular}{|c|c|c|}
+				\hline
+				\textbf{Number} & \textbf{ASCII Bytes} & \textbf{Binary Bytes} \\
+				\hline
+				1               & 1                    & 4                     \\
+				9999            & 4                    & 4                     \\
+				7654321         & 7                    & 4                     \\
+				\hline
+			\end{tabular}
+		\end{center}
+	\end{table}
 \end{frame}
@@
 		\begin{column}{0.5\textwidth}
-			\note[item]{XML example where a huge percent of the actual data sent is metadata}
 			\begin{lstlisting}[language=XML]
@@
 	\end{columns}
-\end{frame}
-
-\begin{frame}[fragile]{Plain Text Example}
-	\begin{table}
-		\caption{Sending A 32-bit Unsigned Integer}
-		\begin{center}
-			\begin{tabular}{|c|c|c|}
-				\hline
-				\textbf{Number} & \textbf{ASCII Bytes} & \textbf{Binary Bytes} \\
-				\hline
-				1               & 1                    & 4                     \\
-				9999            & 4                    & 4                     \\
-				7654321         & 7                    & 4                     \\
-				\hline
-			\end{tabular}
-		\end{center}
-	\end{table}
 \end{frame}
@@
 
-\begin{frame}{Improvements For a New Format}
+\begin{frame}{Improvements For A New Format}
 	\begin{itemize}
@@
 		\item Allow sending metadata to be optional, such as if a client already has it.
-		      \note[item]{Keep metadata and data separate, not coupled together}
 		\item Utilize little-endian to eliminate endian conversions on common architectures.
 		\item Prioritize a simple type system.
-		      \note[item]{Hopefully makes the implementation simpler}
 	\end{itemize}
@@
 				\item Boolean
-				      \note[item]{True of False, one byte}
-				      \note[item]{All integers are little-endian}
-				      \note[item]{Integer types are equivalent to the types in stdint.h C headers}
 				\item U8, U16, U32, U64, U128, U256
-				      \note[item]{Unsigned Integer Types}
 				      \begin{itemize}
@@
 				\item I8, I16, I32, I64, I128, I256
-				      \note[item]{Signed Integer Types}
 				      \begin{itemize}
@@
 				\item F32, F64
-				      \note[item]{Floating Point Types}
 				      \begin{itemize}
@@
 				\item Bytes  (Non-UTF-8 Bytes)
-				      \note[item]{String but doesn't have to be UTF-8}
 				\item String (UTF-8 Bytes)
-				      \note[item]{Sends length as u64 followed by UTF-8 bytes}
 			\end{itemize}
@@
 		\item Next, number of fields as an unsigned 16-bit integer.
-		      \note[item]{A 16-bit integer is used because it's very unlikely for a struct of more than 65 thousand fields to be necessary.}
 		\item Last, pairs of field names and types.
@@
 	\begin{itemize}
-		\item Written in Rust.
-		      \note[item]{systems level programming language similar to C and C++, strong memory safety guarantees}
-		\item Extensively Tested.
-		      \note[item]{100 percent code coverage}
-		      \note[item]{not going to talk a ton about the reference implementation}
-		\item Majority of time was spent on this implementation.
+		\item Written in Rust
+		      \begin{itemize}
+			      \item Efficient code generation similar to C and C++
+			      \item Built-in tagged unions
+			      \item Strong memory safety guarantees
+		      \end{itemize}
+		\item Extensively tested
+		      \begin{itemize}
+			      \item Unit and integration tests
+			      \item 100 percent code coverage
+		      \end{itemize}
 	\end{itemize}
@@
 		\begin{column}{0.5\textwidth}
-			Server
+			Multi-threaded Server
 			\begin{itemize}
 				\item Downloads and parses 1 year of Sony stock history.
-				\item 
+				\item Records the original CSV file size in bytes.
+				\item Records the size in bytes when stored natively in memory.
+				\item Waits to receive requests.
 			\end{itemize}
@@
 		\begin{column}{0.5\textwidth}
-				\item For each data format the client:
+			Client
+			\begin{itemize}
+				\item Formats: CSV, MessagePack, CBOR, JSON, XML, XBF
+				\item For each data format:
 				      \begin{itemize}
@@
 				      \end{itemize}
+			\end{itemize}
 		\end{column}
@@
 
-\section{Results}
+\section{Results and Conclusions}
 
-\begin{frame}{Average Time and Bytes Read}
-	\center
+\begin{frame}[fragile]{Data}
+	\begin{center}
 		\begin{tabular}{|c|c|c|c|c|}
@@
 		\end{tabular}
+	\end{center}
+	Original CSV data size recorded by the server: 17,160 bytes.
+
+	Native data size recorded by the server: 14,558 bytes.
 \end{frame}
 
-\begin{frame}{Overhead}
-	\center
-	\begin{tabular}{|c|c|c|}
-		\hline
-		\textbf{Format} & \textbf{Overhead (bytes)} & \textbf{Percent Overhead} \\
-		\hline
-		CSV             & 1,823                     & 11.1                      \\
-		MessagePack     & 977                       & 6.22                      \\
-		CBOR            & 10,919                    & 42.8                      \\
-		JSON            & 16,592                    & 53.2                      \\
-		XML             & 29,111                    & 66.6                      \\
-		XBF             & 98                        & 0.67                      \\
-		\hline
-	\end{tabular}
+\begin{frame}{CSV Size Difference}
+	CSV Size Difference
+	\begin{itemize}
+		\item Original data always used 8 digits of precision.
+		\item CSV library used only writes the required digits of precision.
+		\item Example: 83.500000 becomes 83.5
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Time Performance}
+	\begin{itemize}
+		\item XBF was faster than every format except MessagePack.
+		\item Formats that operated in binary were faster, even when sending more data (CBOR).
+		\item XBF reference implementation was optimized for memory usage, more could be done for speed.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Size Performance}
+	\begin{itemize}
+		\item XBF required the least amount of bytes.
+		\item Indicates metadata deduplication worked as intended.
+		\item XBF can be even smaller with optimization techniques from MessagePack
+		      \begin{itemize}
+			      \item Coercion to smaller integer sizes when possible.
+			      \item Differently sized variable length types.
+		      \end{itemize}
+		\item Could be possible to have a "negative" overhead.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Use Cases}
+	\begin{itemize}
+		\item Internet of Things and other low bandwidth scenarios.
+		\item High volume applications.
+	\end{itemize}
 \end{frame}
